---
title: "03_Fix_station_duplicates2"
author: "DHJ"
date: "02 09 2025"
output: 
  html_document:
    keep_md: true  
    toc: true
    toc_float: true  
---

* Check (not so much fix) stations that apparently are different, typically slight differences in coordinates between EMODnet and ICES data, as well as differences in "Originator"       
* Calculates the 'duplicate check' dataset 'df_check_all' for all 'station pairs' with the same station name (SD_StationName)     
* Finland not done yet (no values for SD_StationName, needs to rely on coordinates)  

```{r}

knitr::opts_chunk$set(results = "hold")  

```


## 1. Packages    
```{r}

library(dplyr)
library(purrr)
library(furrr)      # future_map
library(tidyr)
library(ggplot2)
library(ggeasy)
library(lubridate)
library(forcats)

source("02_Fix_station_duplicates_functions.R")
last_year <- 2023
```

## 2. Read data  
```{r}

dat_a <- readRDS("Data/01_Combined_data_with_duplicates_a_2025.rds")

cat("nrow(dat_a): ", nrow(dat_a), "\n")

# dat_b <- readRDS("Data/01_Combined_data_with_duplicates_b_2025.rds")

```

### A little cleaning  
```{r}

latlong <- "42.613_14.166"   # Italy
basis <- "W"

dat_a %>%
  filter(LatLong == latlong & BASIS == basis) %>% 
  xtabs(~PARAM + Year, .)

dat_a %>%
  filter(LatLong == latlong & PARAM == "HG" & BASIS == basis) %>% 
  ggplot(aes(Year, Value, color = Dataset, size = Dataset)) +
  geom_point() +
  scale_size_manual(values = c(2,1)) +     # make 'val1' dots bigger
  easy_rotate_labels()

sel <- with(dat_a, LatLong == latlong & BASIS == basis & Value > 1000); sum(sel)  
dat_a <- dat_a[!sel,]

```
#### Fix fat weight  
```{r}

if (FALSE)
  table(addNA(dat_a$FATWT_MUNIT))
#       %     g/g    <NA> 
#  241142     257 1312129 

sel <- dat_a$FATWT_MUNIT %in% "g/g"
dat_a$FATWT[sel] <- dat_a$FATWT[sel]*100
dat_a$FATWT_MUNIT[sel] <- "%"
cat("Changed", sum(sel), "fat fraction unit to % \n\n")

cat("Fat weight fraction, unit: \n")
table(addNA(dat_a$FATWT_MUNIT))
cat("\n\n")
cat("Dry weight fraction, unit: \n")
table(addNA(dat_a$DRYWT_MUNIT))

  
```


## 3. Check duplication by station   
* In some cases, data which apparently comes from the same staton the same year (with same measurement values)
* In other cases, there are not data from the same year, but times series are fragmented because they apparently belong to different stations   

### Most common variables   
- Used in duplicate check below  
```{r}

tab_loq_mean <- dat_a %>%
   group_by(PARAM) %>% 
   summarise(over_LOQ = mean(is.na(QFLAG))) %>% 
   arrange(desc(over_LOQ)) 

# pull(tab_loq_mean, PARAM) %>% head(30)

tab_loq_n <- dat_a %>%
   group_by(PARAM) %>% 
   summarise(over_LOQ = sum(is.na(QFLAG))) %>% 
   arrange(desc(over_LOQ)) 

# pull(tab_loq_n, PARAM) %>% head(30) %>% dput()

# list of 20
param_selected <- c(
  "HG", "CD", "ZN", "CU", "PB", "NI", "CR", 
  "CB153", "CB118", "CB138", "CB101", 
  "DDEPP", "HCB", "HCHG", "HCHA","BDE47", 
  "FLU", "PYR", "PA", "BAP")
length(param_selected)

# list of 13
param_selected <- c(
  "HG", "CD", "ZN", "CU", "PB",  
  "CB153", "CB118",  
  "DDEPP", "HCB", "HCHG", "BDE47", 
  "FLU", "BAP")
length(param_selected)

# List of 18, including all 7 PCBs
param_selected <- c(
  "HG", "CD", "ZN", "CU", "PB",  
  "CB28", "CB52", "CB101", "CB118", "CB138", "CB153", "CB180",
  "DDEPP", "HCB", "HCHG", "BDE47", 
  "FLU", "BAP", "PFOS", "PFAS")
length(param_selected)


```
### Species-specific bivalve fat and dry weights  
- From OSPAR (https://ocean.ices.dk/ohat/trDocuments/2021/help_ac_basis_conversion.html )   
- For MAcoma baltica, see part 8  
```{r}

df_bivalve_dryweight <- read.csv(textConnection("
WoRMS_scientificName, Common_name, FATWT_species, DRYWT_species
Cerastoderma edule,	common cockle, NA, 19.0 		
Mya arenaria, softshell clam, 0.7, 14.8 		
Ruditapes philippinarum, manila clam, NA, 16.0 		
Mytilus edulis, blue mussel, 1.4, 16.3 		
Mytilus galloprovincialis, Mediteranean mussel, 2.2, 19.0 		
Crassostrea gigas, Pacific oyster, 2.1, 18.0 		
Ostrea edulis, native oyster, 1.8, 20.4,
Macoma balthica, Baltic clam, NA, 19.96,
"), 
stringsAsFactors = FALSE,
colClasses = c("character", "character", "double", "double"))

saveRDS(df_bivalve_dryweight, "Data/03_df_bivalve_dryweight_2025.rds")   # should be changed to 2024?

species_selected <- df_bivalve_dryweight$WoRMS_scientificName

# Check
tab <- table(dat_a$WoRMS_scientificName)
cat(100*mean(species_selected %in% rownames(tab)), 
    "% of names found in dat_a \n")

dat_a %>%
  filter(WoRMS_scientificName %in% species_selected) %>%
  count(WoRMS_scientificName)

tab[order(tab)] %>% rev() %>% head(30)

```


### Matrix used for bivalves   
```{r}

# Check matrices used  
#
# FA, MU&FA and SB  
# SB = whole soft body
# MU&FA = muscle and fat  
# FA = fat, seems to be used as a very uncommon "synonym" of BASIS = "F" (sometimes as dublettes)
# TM = tail muscle, used for a very few Mytilus edulis (!)
# WO = whole organism, used for a very few Mytilus edulis  
# MU used for PCB in two years in Lithuania and United Kingdom (for both, SD_StationName = NA; in UK 3-4 locations)

tab <- dat_a %>%
  filter(PARAM %in% param_selected & WoRMS_scientificName %in% species_selected) %>% 
  xtabs(~MATRX, .)
cat("Numbers by matrix: \n")
tab

cat("\n\nPercentages: \n")
round(tab/sum(tab)*100, 2)

cat("\n\nPercentage MU&FA + SB: \n")
round(sum(tab[c("MU&FA", "SB")])/sum(tab)*100, 2)
# 99.7% of t

# MATRX
# FA     MU  MU&FA     SB     TM     WO 
# 16     59 110673 193230     16     14 


cat("\n\nNumbers by species: \n")
dat_a %>%
  filter(PARAM %in% param_selected & WoRMS_scientificName %in% species_selected) %>%
  xtabs(~WoRMS_scientificName  + MATRX, .)

if (FALSE){
  
  # example of "FA" 
  df <- dat_a %>%
    filter(grepl("CB", PARAM) & SD_StationName == "Byttelocket" & WoRMS_scientificName %in% "Mytilus edulis" & Year == 2006)
  xtabs(~PARAM + MATRX, df)
  xtabs(~BASIS + MATRX, df)
  
  dat_a %>%
    filter(PARAM %in% param_selected & WoRMS_scientificName %in% species_selected & MATRX %in% "MU") %>%
    xtabs(~addNA(SD_StationName), .)

  # MATRX %in% "MU"
  df <- dat_a %>%
    filter(PARAM %in% param_selected & WoRMS_scientificName %in% species_selected & MATRX %in% "MU")
  xtabs(~Year + addNA(SD_StationName) + Country, df)
  View(df)
  

}

```
## 4a. Select subset of data   

### dat_2: Subset by species, matrix and parameters   
-From now on, we can ignore matrix  
```{r}

dat_2 <- dat_a %>%
  filter(
    WoRMS_scientificName %in% species_selected &
      MATRX %in% c("MU&FA", "SB") &
      PARAM %in% param_selected
      )

nrow(dat_a)
nrow(dat_2)

```

### Units   
Fixed already  
```{r}

dat_2 %>%
  xtabs(~addNA(UNIT), .)

```

### Check duplicates within ICES data  
- Same SampleIdentifier, SubsampleIdentifier (so we only check ICES, these are empty for EMODnet)  
- Check 
    - Some observations recorded with two different units    
    - Most have just two observations of the same unit  
    - Difference can be quite big   
    - Very often one below detection linit, one above    
```{r}

if (FALSE){
  # example
  dat_2 %>%
    filter(grepl("Porto de Pescas", SD_StationName) & Year == 2010 & PARAM == "CB118") %>%
    select(SampleIdentifier, SubsampleIdentifier, PARAM, 
           Value_orig, UNIT_orig, Value, UNIT)
}

  
df_dublettes <- dat_2 %>%
  filter(Dataset == "ICES") %>%
  group_by(SampleIdentifier, SubsampleIdentifier, PARAM) %>%
  arrange(SampleIdentifier, SubsampleIdentifier, PARAM, UNIT_orig) %>%
  mutate(n = n(), 
         Units = paste(unique(UNIT_orig), collapse = ","),
         Mean = mean(Value),
         Range = diff(range(Value)),
         Range_rel = Range/Mean) %>%
  ungroup()
         

xtabs(~n, df_dublettes)

df_dublettes %>%
  filter(n > 1) %>%
  xtabs(~Units, .)

if (FALSE){
  
  df_dublettes %>%
    filter(n > 1 & Units == "ug/g") %>%
    select(SD_StationName, SD_StationCode, Year, Samplingtime, SampleIdentifier,
           Value, UNIT, Units, Mean, Range, Range_rel) %>% 
    View()
  
}

#
# Dublettes with different original units
#
df <- df_dublettes %>%
  filter(n > 1 & Units %in% c("ng/g,ng/kg"))

if (FALSE){
  df %>%
    select(SD_StationName, SD_StationCode, Year, Samplingtime, SampleIdentifier,
           Value, QFLAG,UNIT_orig, UNIT, Mean, Range, Range_rel) %>%
    View()
}

#
# Dublettes with same original unit
#
df <- df_dublettes %>%
  filter(n > 1 & Units %in% c("ug/g","ug/kg")) 

if (FALSE){
  df %>%
    select(SD_StationName, SD_StationCode, Year, Samplingtime, SampleIdentifier,
           Value, QFLAG, UNIT_orig, UNIT, Mean, Range, Range_rel) %>%
    View()
}



```
### dat_3: remove ICES duplicates    
- We pick the highest one (and it's QFLAG value)   
```{r}

data_for_replacement <- dat_2 %>%
  filter(Dataset == "ICES") %>%
  group_by(SampleIdentifier, SubsampleIdentifier, PARAM) %>%
  arrange(SampleIdentifier, SubsampleIdentifier, PARAM, desc(Value)) %>%
  mutate(n = n(), 
         value_no = 1:length(Value),
         Value = first(Value),
         QFLAG = first(QFLAG),
         .groups = "drop") %>%
  ungroup() %>%
  filter(
    n > 1 &           # pick duplicates 
    value_no == 1)    # pick no. 1 of the duplicates
cat(nrow(data_for_replacement), "rows of data to replace duplicates")

dat_3_emodnet <- dat_2 %>%
  filter(Dataset == "EMODnet")

dat_3_ices_without_dupl <- dat_2 %>%
  filter(Dataset == "ICES") %>%
  anti_join(
    data_for_replacement, by = c("SampleIdentifier", "SubsampleIdentifier", "PARAM")
  )

n1 <- nrow(dat_2 %>% filter(Dataset == "ICES"))
n2 <- nrow(dat_3_ices_without_dupl)
cat("\n")
cat(n1-n2, "duplicate rows removed from dat_2 \n")
cat("'data_for_replacement' can replace", nrow(data_for_replacement)*2, "rows \n")
cat("  (as all duplicates are 2 copies) \n")

dat_3 <- bind_rows(
  dat_3_emodnet,
  dat_3_ices_without_dupl,
  data_for_replacement %>% select(-n, -value_no)
)

n1 <- nrow(dat_2)
n2 <- nrow(dat_3)

cat("\n")
cat("'dat_2' had", nrow(dat_2), "rows \n")
cat("'dat_3' has", nrow(dat_3), "rows \n")
cat("- difference =", n1-n2, "rows \n")



```

### Table: Number of years per country/dataset and parameter   
* Note that some countries, (Denmark, Estonia, Finland, Sweden) have time series from both ICES and EMODnet  
* Means possible duplication  
```{r}

#  [1] "CB138" "CB153" "HCHG"  "TDEPP" "CB180" "PB"    "CB105" "CB28"  "CD"    "CB101" "CB52" 
# [12] "BKF"   "BGHIP" "ANT"   "BAP"   "CR"    "FLE"   "CB118" "CU"    "DDTPP" "CB156" "ACNLE"
# [23] "HG"    "DDEPP" "NAP"   "PA"    "ZN"    "BBF"   "FLU"   "BAA"   "NI"    "ACNE"  "DBAHA"
# [34] "CHR"   "ICDP"  "PYR"   "TBTIN" "HCB"   "HCHA"  "ALD"   "LNMEA" "AS"    "HCBD"  "DIELD"
# [45] "CB128" "CB194" "HCHB"  "TBSN+" "CB170" "DBT"   "CB183" "BDE47" "BD100" "BD153" "BDE99"
# [56] "BD154" "SCCP"  "DEHP"  "TPSN+" "BBJKF" "CHRTR" "HCEPX" "ISOD"  "HEPC"  "ENDA"  "END"  
# [67] "QCB" 

# Number of years per country/dataset for a selectiopn of parameters  
tab_a <- dat_3 %>%
  filter(PARAM %in% param_selected) %>%
  mutate(Country = paste0(Country, "_", substr(Dataset, 1, 4))) %>%
  distinct(PARAM, Country, Year) %>%
  xtabs(~Country + PARAM, .)  

tab_a

# tab_b <- dat_b %>%
#   filter(PARAM %in% c("CD","HG","PB","HCB","HCHG","DDEpp","CB101","BAP")) %>%
#   mutate(Country = paste0(Country, "_", substr(Dataset, 1, 4))) %>%
#   distinct(PARAM, Country, Year) %>%
#   xtabs(~Country + PARAM, .)  
# 
# tab_diff <- tab_a - tab_b
  
# colnames(tab_a) <- paste0(colnames(tab_a), "_a")
# colnames(tab_b) <- paste0(colnames(tab_b), "_b")

# tab_comb <- cbind(tab_a, tab_b)  
# tab_comb

```
### Check SD_StationCode  
- Only in 3 cases does one SD_StationName have >1  SD_StationCode  
- No iverlap in years, probably just a minor change in coordinates  
- We assume they are the same (i.e, we use 'SD_StationName')   
```{r}

df1 <- dat_3 %>% 
  filter(!is.na(SD_StationName)) %>%
  distinct(SD_StationName, SD_StationCode) %>%
  count(SD_StationName) %>%
  filter(n > 1)
# nrow(df1)
df1

dat_3 %>%
  filter(SD_StationName %in% pull(df1, SD_StationName)) %>%  
  xtabs(~Year + SD_StationCode + SD_StationName, .)  

cat("Max distances (km) for the 3 stations with different 'SD_StationCode': \n")
i <- 1
df2 <- dat_3 %>%
  filter(SD_StationName %in% pull(df1, SD_StationName)[i]) %>%  
  distinct(SD_StationCode, Longitude, Latitude) %>%
  mutate(id = paste(SD_StationCode, Longitude, Latitude, sep = "_"))  
df3 <- earth.dist.pairwise(df2, var_id = "id")
pull(df3, dist) %>% max(na.rm = TRUE)
# < 250 m distance. Older data (2005) and changed SD_StationCode in 2004/05  

i <- 2
df2 <- dat_3 %>%
  filter(SD_StationName %in% pull(df1, SD_StationName)[i]) %>%  
  distinct(SD_StationCode, Longitude, Latitude) %>%
  mutate(id = paste(SD_StationCode, Longitude, Latitude, sep = "_"))
df3 <- earth.dist.pairwise(df2, var_id = "id")
pull(df3, dist) %>% max(na.rm = TRUE)
# 1.3 km distance. Newer data (2018) but changed SD_StationCode in 2001  

i <- 3
df2 <- dat_3 %>%
  filter(SD_StationName %in% pull(df1, SD_StationName)[i]) %>%  
  distinct(SD_StationCode, Longitude, Latitude) %>%
  mutate(id = paste(SD_StationCode, Longitude, Latitude, sep = "_"))
df3 <- earth.dist.pairwise(df2, var_id = "id")
pull(df3, dist) %>% max(na.rm = TRUE)
# 3.1 km distance. Old data, up to 2001  

```


### Add ID  
```{r}

# xtabs(~is.na(LatLong) + is.na(SD_StationName), dat_3)

dat_3 <- dat_3 %>%
  mutate(
    id = case_when(
      !is.na(SD_StationName) ~ SD_StationName,
      is.na(SD_StationName) ~ LatLong
    )
  )

```

### Check SubsampleIdentifier 
- Are there >1 per SampleIdentifier?  
```{r}

df1 <- dat_3 %>% 
  filter(!is.na(SampleIdentifier)) %>%
  distinct(SampleIdentifier, SubsampleIdentifier) %>%
  count(SampleIdentifier) %>%
  filter(n > 1)
# nrow(df1)
table(df1$n)


```
### dat_4: EMODnet data summarized  
- Summarize per station/year  
- We need to do that to summarize PCB7, since we have no way to distinguish samples    
```{r}

# dat_3 %>% names() %>% paste(collapse = ", ")

dat_4_emod <- dat_3 %>%
  filter(Dataset == "EMODnet") %>%
  group_by(
    Dataset, Year, Month, MSFD_region, Country, SD_StationName, MYEAR, Originator,
    Samplingtime, Latitude, Longitude, SampleIdentifier, SubsampleIdentifier,
    WoRMS_scientificName, WoRMS_AphiaID, WoRMS_accepted_scientificName, WoRMS_accepted_AphiaID, 
    SEXCO, PARAM, PARGROUP, MATRX, BASIS, id) %>%
  summarise(
    Value = median(Value, na.rm = TRUE),
    QFLAG_mean = mean(!is.na(QFLAG)),
    QFLAG = ifelse(QFLAG_mean > 0.5, "<", as.character(NA))
    )

dat_4 <- bind_rows( 
  dat_3 %>% filter(Dataset == "ICES"),
  dat_4_emod %>% select(-QFLAG_mean)
  )

```




## 4b. Sum PCB   

### Definitions  

```{r}

param_pcb7 <- c("CB28", "CB52", "CB101", "CB118", "CB138", "CB153", "CB180")
param_pcb6 <- c("CB28", "CB52", "CB101", "CB138", "CB153", "CB180")

```


### Check data  

```{r, fig.width=9, fig.height=4}

dat_4 %>%
  filter(PARAM %in% param_pcb7) %>%
  xtabs(~BASIS + Dataset, .)

dat_4 %>%
  filter(PARAM %in% param_pcb7) %>%
  xtabs(~BASIS + Country, .)

dat_4 %>%
  filter(PARAM %in% param_pcb7 & BASIS == "W" & Year >= 2010) %>%
  mutate(PARAM = factor(PARAM, levels = param_pcb7)) %>%
  group_by(id, PARAM) %>%
  summarise(Value_median = median(Value)) %>%
  ggplot(aes(id, PARAM, fill = Value_median)) +
  geom_tile() +
  easy_rotate_x_labels()

```


### PCBs on broad format     
```{r}

df_pcb_1 <- dat_4 %>%
  filter(PARAM %in% param_pcb7) %>%
  mutate(PARAM = factor(PARAM, levels = param_pcb7)) %>%
  select(Dataset, Year, Month, MSFD_region, Country, SD_StationName, MYEAR, Originator,
    Samplingtime, Latitude, Longitude, SampleIdentifier, SubsampleIdentifier,
    WoRMS_scientificName, WoRMS_AphiaID, WoRMS_accepted_scientificName, WoRMS_accepted_AphiaID, 
    SEXCO, PARAM, PARGROUP, MATRX, BASIS, id, Value)


df_pcb <- df_pcb_1 %>%
  pivot_wider(names_from = PARAM, values_from = Value)

# if you get 'Warning: Values are not uniquely identified; output will contain list-cols'
if (FALSE)
  check <- df_pcb_1 %>%
    pivot_wider(names_from = PARAM, values_from = Value, values_fn = length)

# Sum PCB7
df_pcb$PCB7 <- apply(df_pcb[param_pcb7], 1, sum, na.rm = TRUE)
# Number of PCBs
df_pcb$PCB7_n <- apply(!is.na(df_pcb[param_pcb7]), 1, sum)

# Sum PCB6
df_pcb$PCB6 <- apply(df_pcb[param_pcb6], 1, sum, na.rm = TRUE)
# Number of PCBs
df_pcb$PCB6_n <- apply(!is.na(df_pcb[param_pcb6]), 1, sum)

# Table of number of PCBs
tab <- xtabs(~Year + PCB7_n, df_pcb)
tab <- cbind(tab, apply(tab, 1, sum))

# Plot 
plot(as.numeric(rownames(tab)), 100*tab[,7]/tab[,8], 
     type = "b",
     main = "Percentage with 7 PCBs", ylab = "Percent")

# Plot of PCB7 vs PCB6  
ggplot(df_pcb, aes(PCB6, PCB7)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1)

df_pcb %>%
  mutate(Ratio = PCB7/PCB6) %>%
  ggplot(aes(PCB6, PCB7, color = Ratio)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1)

if (FALSE){
  
  stem(df_pcb$PCB7 / df_pcb$PCB6)

  # Some stations have extremely high CB118 values compare to rest of CBs 
  df_pcb %>%
    mutate(Ratio = PCB7/PCB6) %>%
    filter(Ratio > 2) %>%
    View("Hi PCB7/PCB6")
  
  
}


```


### Flag for PCB6 + PCB7
```{r}

df_pcb_2 <- dat_4 %>%
  filter(PARAM %in% param_pcb7) %>%
  mutate(PARAM = factor(PARAM, levels = param_pcb7),
         QFLAG = ifelse(is.na(QFLAG), as.character(NA), "<")) %>%
  select(Dataset, Year, Month, MSFD_region, Country, SD_StationName, MYEAR, Originator,
    Samplingtime, Latitude, Longitude, SampleIdentifier, SubsampleIdentifier,
    WoRMS_scientificName, WoRMS_AphiaID, WoRMS_accepted_scientificName, WoRMS_accepted_AphiaID, 
    SEXCO, PARAM, PARGROUP, MATRX, BASIS, id, QFLAG)

df_pcb7_flag <- df_pcb_2 %>%
  pivot_wider(names_from = PARAM, values_from = QFLAG)

df_pcb6_flag <- df_pcb_2 %>%
  filter(PARAM %in% param_pcb6) %>%
  pivot_wider(names_from = PARAM, values_from = QFLAG)

# Fraction of flags indicating <LOQ 
df_pcb7_flag$QFLAG_mean <- apply(!is.na(df_pcb7_flag[param_pcb7]), 1, mean)
df_pcb6_flag$QFLAG_mean <- apply(!is.na(df_pcb6_flag[param_pcb6]), 1, mean)

# New flag
df_pcb7_flag$QFLAG <- ifelse(df_pcb7_flag$QFLAG_mean > 0.5, "<", as.character(NA))
df_pcb6_flag$QFLAG <- ifelse(df_pcb6_flag$QFLAG_mean > 0.5, "<", as.character(NA))

df_pcb7_flag %>%
  group_by(Year) %>%
  summarise(QFLAG_mean = mean(QFLAG_mean)) %>%
  ggplot(aes(Year, QFLAG_mean)) + geom_point() +
  labs(main = "Flag of PCB7")

df_pcb6_flag %>%
  group_by(Year) %>%
  summarise(QFLAG_fraction = mean(!is.na(QFLAG))) %>%
  ggplot(aes(Year, QFLAG_fraction)) + geom_point() +
  labs(main = "Flag of PCB6")

```


### dat_5: Add PCB7 + PCB6 to data  

```{r}

df_pcb7_values <- df_pcb %>%
  select(Dataset:id, PCB7) %>%
  rename(Value = PCB7)
nrow(df_pcb7_values)

df_pcb6_values <- df_pcb %>%
  select(Dataset:id, PCB6) %>%
  rename(Value = PCB6)
nrow(df_pcb6_values)

df_pcb7_for_adding <- df_pcb7_values %>%
  mutate(
    PARAM = "PCB7",
    QFLAG = df_pcb7_flag$QFLAG)

common_names <- c(
  c("Dataset", "Year", "Month", "MSFD_region", "Country", "SD_StationName", "MYEAR", 
    "Originator", "Samplingtime", "Latitude", "Longitude", "SampleIdentifier", 
    "SubsampleIdentifier", "WoRMS_scientificName", "WoRMS_AphiaID", "WoRMS_accepted_scientificName",
    "WoRMS_accepted_AphiaID", "SEXCO", "PARGROUP", "MATRX", "BASIS", "id"))

df_pcb6_for_adding <- df_pcb6_values %>%
  left_join(df_pcb6_flag[c(common_names, "QFLAG")],
            by = common_names) %>%
  mutate(
    PARAM = "PCB6")

nrow(df_pcb7_for_adding)
nrow(df_pcb6_for_adding)

dat_5 <- bind_rows(
  dat_4,
  df_pcb7_for_adding,
  df_pcb6_for_adding)  

```

### Percentage PCB7 over LOQ     
```{r}

dat_5 %>%
  filter(WoRMS_scientificName %in% species_selected) %>%
  filter(PARAM %in% c(param_pcb7, "PCB7")) %>%
  mutate(PARAM = factor(PARAM, levels = c(param_pcb7, "PCB7"))) %>%
  group_by(PARAM, Year) %>%
  summarise(over_loq_perc = round(100*mean(is.na(QFLAG)),1),
            .group = "drop") %>%
  filter(Year >= 2000) %>%
  ggplot(aes(Year, over_loq_perc, color = PARAM)) +
  geom_line()

```

## 5. Select by month


### Distribution of month by country  
- Note early spring data in UK (esp. Feb-March)    
- Note that Croatia has mainly March data   
- Big spread 
```{r}

if (FALSE)
  dat_5 %>% xtabs(~Country + Month, .)

dat_5 %>%
  ggplot(aes(Month)) +
  geom_histogram(binwidth = 1) +
  facet_wrap(vars(Country), scales = "free_y")

```

### Examples - one country/parameter/year by month    
- The source of discrepancies between ICES and EMODnet data is that EMODnet also contains spring data    
```{r, fig.width=9, fig.height=6}

param <- "HG"

country <- "France"; year <- 2014; basis <- "D"
# country <- "Italy"; year <- 2017; basis <- "W"
# country <- "Germany"; year <- 2015; basis <- "D"
# country <- "Ireland"; year <- 2018; basis <- "W"
# country <- "Ukraine"; year <- 2005; basis <- "W"
# country <- "Slovenia"; year <- 2005; basis <- "W"
# country <- "The Netherlands"; year <- 2019; basis <- "W"

# For Slovenia:
# dat_3 %>%
#   filter(Country == country) %>%
#   xtabs(~PARAM + Year + BASIS, .)

dat_3 %>% 
  filter(Country == country & PARAM == param) %>%
  xtabs(~BASIS + Year, .)

# Ignoring month     
dat_3 %>% 
  filter(Country == country & PARAM == param & Year == year & BASIS == basis) %>% 
  # View()
  ggplot(aes(id, Value, color = Dataset, size = Dataset)) +
  geom_point() +
  scale_size_manual(values = c(2,1)) +     # make 'val1' dots bigger
  easy_rotate_labels()

# Faceted by month 
dat_3 %>% 
  filter(Country == country & PARAM == param & Year == year & BASIS == basis) %>%
  ggplot(aes(id, Value, color = Dataset, size = Dataset)) +
  geom_point() +
  scale_size_manual(values = c(2,1)) +    # make 'val1' dots bigger
  facet_wrap(vars(Month)) +
  easy_rotate_labels()

```

#### UK  
```{r, fig.width=9, fig.height=6}

if (FALSE){
  dat_5 %>% 
    filter(Country == "United Kingdom") %>% xtabs(~Year + PARAM + BASIS, .)
}


# Based on the raw data - faceted by month 
dat_5 %>% 
  filter(Country == "United Kingdom" & PARAM == "HG" & Year == 2018 & BASIS == "W") %>%
  ggplot(aes(id, Value, color = Dataset, size = Dataset)) +
  geom_point() +
  scale_size_manual(values = c(2,1)) +    # make 'val1' dots bigger
  facet_wrap(vars(Month)) +
  easy_rotate_labels("x")

```


### dat_6: Delete French spring data  
```{r}

dat_6 <- dat_5 %>%
  filter(!(Country == "France" & Month <= 6))

nrow(dat_5)
nrow(dat_6)

```

### Save dat_6  
```{r}
  
saveRDS(dat_6, "Data/03_dat_6_2025.rds")

```


## 6. Basis conversion    
- Add species-specific DW and fat weight to data  
- Dry weight: Use sample-specific when possible, otherwise species-specific  
- Fat weight: Use species-specific anyway  
```{r}

table(!is.na(dat_6$DRYWT), addNA(dat_6$DRYWT_MUNIT))

dat_7 <- dat_6 %>%
  left_join(df_bivalve_dryweight) %>%
  mutate(
    DRYWT = case_when(
      !is.na(DRYWT) & DRYWT_MUNIT %in% "%" ~ DRYWT,  # some 1000s lack 'DRYWT_MUNIT'  
      TRUE ~ DRYWT_species
    ),
    DRYWT_MUNIT = "%"
    )

# table(!is.na(dat_7$DRYWT), addNA(dat_7$DRYWT_MUNIT))
# nrow(dat_6)
cat("nrow(dat_7): ", nrow(dat_7), "\n")

```

### Save dat_7    

* These are the main data used onwards in script 04 
- Note: we do *not* filter by temporal coverage at this point, as ICES + EMODnet time series must by joined first   

```{r}

saveRDS(dat_7, "Data/03_dat_7_2025.rds")

```

## 7a. Duplicate check, same name         

- Duplicate check of all stations with same name   
- Note: when reading tables/graphs, "1" is EMODnet and "2" is ICES (e.g. val1, val2)  

#### - stations in both EMODnet and ICES data    
Table shows number of years per station and dataset  
```{r}

param_selected_cb7 <- c(param_selected, "PCB7")

table(dat_7$Dataset)

tab_a <- dat_7 %>%
  filter(PARAM %in% param_selected_cb7) %>%
  mutate(Country = paste0(Country, "_", substr(Dataset, 1, 4))) %>%
  distinct(id, Country, Dataset, Year) %>% # View()
  xtabs(~id + Dataset, .)  

cat("Dataset with EMODnet data a \n\n")
sel <- apply(tab_a>0, 1, sum) > 1
cat(sum(sel), "stations have data in both datasets\n")  
cat(100*round(mean(sel),3), "% of stations have data in both datasets\n\n")  

# Get stations with data in both datasets
stations_dupl <- rownames(tab_a[sel,])

```

#### - prepare data sets  

```{r}

# Make data set with extra stations for checking (see 'DK4' in script 1)
dat_test <- dat_7 %>%
  filter(id %in% stations_dupl) %>%
  filter(PARAM %in% param_selected_cb7) %>%
  mutate(
    Station_dataset = paste0(id, "_", Dataset),  # 'station' names  
    MATRX = ifelse(MATRX %in% "MU&FA", "SB", MATRX)          # hypothesis: these two are the same
  )

# distance between ICES and EMODnet stations - will be added later  
dat_distance <- dat_test %>%
  group_by(id, Dataset) %>%
  summarise(across(c(Longitude, Latitude), first), .groups = "drop") %>%
  ungroup() %>%
  pivot_wider(names_from = Dataset, values_from = c(Longitude, Latitude)) %>%
  # calculating distance using 'earth.dist'  
  mutate(distance = earth.dist(
    Longitude_EMODnet, Latitude_EMODnet, 
    Longitude_ICES, Latitude_ICES)
  )

# debugonce(check_one_pair_by_stationname)
# test <- check_one_pair_by_stationname(stations_dupl_sel[1], dat_test)

```




#### - positions  
```{r}

dat_test_pos <- dat_test %>%
  group_by(id, Dataset) %>%
  summarise(across(c(Longitude, Latitude), first), .groups = "drop")

```


#### - test  
```{r}

if (FALSE){
  
  debugonce(check_pair_by_stationname)
  debugonce(check_pair)
  debugonce(check_pair_matrix_param_year)
  test <- check_pair_by_stationname("Aber Benoît", data = dat_test)
  
}

```

#### - make 'df_check_all'   
This takes some minutes  
```{r}

# stations_dupl <- stations_dupl[1:5]

df_check_all_filename <- "Data/03_df_check_all_1a_2025.rds"

# If we want to perform analysis anyway, set to TRUE:
recompute_df_check_all <- FALSE

#
# Perform analysis only if there is no saved result 
#   (or 'recompute_df_check_all' = TRUE)
#
if (!file.exists(df_check_all_filename) | recompute_df_check_all){
  
  # Check cores
  future::availableCores()
  
  # Set a "plan" for how the code should run.
  plan(multisession, workers = 10)
  
  # 'safe' version of check_pair_by_stationname   
  check_pair_by_stationname_s <- safely(check_pair_by_stationname)  
  
  # Run function (takes some minutes)
  # Note: dataset 1 (val1 etc.) will always be EMODnet, dataet 2 = ICES 
  t0 <- Sys.time()
  check_all_list1 <- stations_dupl %>% 
    future_map(
      check_pair_by_stationname_s, 
      data = dat_test,
      logfile = "03_logfile_check_same_stationname.txt")
  t1 <- Sys.time()
  t1-t0   # 6 mins (19 parameters)
  
  # Transpose and keep only the ones that work
  check_all_list2 <- transpose(check_all_list1) 
  
  # when the 'error' part is NULL, there is no error
  is_ok <- check_all_list2$error %>% map_lgl(is.null)
  message(100*mean(is_ok), " % of stations had no errors")
  
  # Filter list, keeping only those that are ok
  check_all_list3 <- check_all_list2[is_ok]
  
  # Make final check table
  df_check_all_1a <- check_all_list3$result %>% 
    bind_rows()
  
  # Add coordinates plus distance
  n1 <- nrow(df_check_all_1a)
  df_check_all_1a <- df_check_all_1a %>%
    left_join(dat_distance, by = "id")
  n2 <- nrow(df_check_all_1a)
  if (n2 > n1) stop("Number of rows increased")
  
  if (file.exists(df_check_all_filename)){
    
    # Save backup
    timestamp <- Sys.time() %>% gsub(" ", "_", .) %>% gsub(":", "-", .)
    new_filename <- paste0(
      sub(".rds", "", df_check_all_filename, fixed = TRUE),
      "_", timestamp, ".rds"
    )
    file.copy(df_check_all_filename, new_filename)
    
  }
  
  # SAVE
  saveRDS(df_check_all_1a, file = df_check_all_filename)
  
} else {
  
  df_check_all_1a <- readRDS(df_check_all_filename)
  
}

```

#### - check the ones with more EMODnet than OSPAR data   
- very few instances of this after 2010, and only in two stations: LIM-3728-3 and STR 0102065   
```{r}

if (FALSE) {
  
df_check_all_1a %>% 
  xtabs(~n1 + n2, .)

df_check_all_1a %>%
  filter(n2 >= 1 & n1 > n2) %>%
  xtabs(~id + n2, .)  

df_check_all_1a %>%
  filter(n2 >= 1 & n1 > n2 & year > 2010) %>% View()

}


df_check_all_1a %>%
  filter(n2 >= 1 & n1 > n2) %>%
  xtabs(~year, .)  

df_check_all_1a %>%
  filter(n2 >= 1 & n1 > n2 & year > 2010) %>% 
  xtabs(~id + year, .)  


```

## 7b. Filter by temporal coverage   

### Filtering     

- For levels, we need 3 years in the period 2010-2020          
- For 10-year trends, we need at least 5 years in the period 2005-2020 and last year >= 2015    

```{r}

df_check_all_1a <- df_check_all_1a %>%
  group_by(id, distance) %>%
  mutate(n_year = length(unique(year)),
         n_year_2010.20 = length(unique(year[year >= (last_year - 10) & year <= last_year])),
         n_year_2005.20 = length(unique(year[year >= (last_year - 5) & year <= last_year])),
         last_year = max(year),
         levels_ok = n_year_2010.20 >= 3,
         trend10yr_ok = n_year_2005.20 >= 5 & last_year >= 2015) %>%
  ungroup()

df_check_all_1b <- df_check_all_1a %>%
  filter(levels_ok | trend10yr_ok)

cat("df_check_all_1a, no of stations:", length(unique(df_check_all_1a$id)), "\n")
cat("df_check_all_1b, no of stations:", length(unique(df_check_all_1b$id)), "\n")

if (FALSE){
  
  df_check_all_1a %>% filter(id == "MGAVSA") %>% xtabs(~n1 + n2, .)
  df_check_all_1b %>% filter(id == "MGAVSA") %>% xtabs(~n1 + n2, .)
  
}

```


### Check longest distances   
 
#### - manual check    
- dist_longest is also sorted  
```{r, fig.width=11, fig.height = 6}

dist_longest <- df_check_all_1b %>%
  distinct(id, distance, n_year, last_year, n_year_2010.20, param) %>%
  count(id, distance, n_year, last_year, n_year_2010.20, ) %>%
  arrange(desc(distance)) %>%
  as.data.frame() 

check_plot <- function(station_number, param_sel = NULL){
  gg <- plot_pair_by_stationname(
    dist_longest$id[station_number], 
    checkdata = df_check_all_1b,
    param_sel = param_sel)    # MGAVSA
  gg <- gg + labs(subtitle = paste("Distance", dist_longest$distance[station_number]))
  print(gg)
}


dist_longest$id[1]
# debugonce(plot_pair_by_stationname)
check_plot(1)
check_plot(1, c("HG", "DDEPP", "HCB", "HCHG"))
check_plot(1, "DDEPP")
check_plot(2)
check_plot(2, "CB118")

# dist_longest 1 and 3 - Keep ICES only for these: 
# "MGAVSA"	Mytilus edulis
# "KBKLYN"	Mytilus edulis

# The rest (2,4,5,6, etc.) seems OK
# debugonce(plot_pair_by_stationname)

if (FALSE){
  i <- 1
  df_check_all_1b %>% filter(
    id %in% dist_longest$id[i] & param %in% "DDEPP") %>% View(dist_longest$id[i])
  i <- 2
  df_check_all_1b %>% filter(
    id %in% dist_longest$id[i] & param %in% "CB118") %>% View(dist_longest$id[i])
}

```

#### - create 'df_check_all2'  
- Filtered by dist_longest (takes care of amount of data since 2011)  
- Plus two extra stations that are removed 
```{r}

# Exclude these (see previous chunk):
# "MGAVSA"	Mytilus edulis (ICES) + M. galloprovincialis (EMODnet)
# "KBKLYN"	Mytilus edulis
# EMODnet part will be excluded from the main data (see 06)
dist_longest_manually_filtered <- dist_longest %>%
  filter(!id %in% c("MGAVSA", "KBKLYN"))

df_check_all2 <- df_check_all_1b %>%
  left_join(
    dist_longest_manually_filtered %>% 
      select(id, n_year, last_year) %>%
      rename(station_n_year = n_year, station_last_year = last_year),
    by = c("id")
  ) %>%
  filter(!is.na(station_n_year))
  
cat("No of rows, df_check_all_1b:", nrow(df_check_all_1b), "\n")  
cat("No of rows, df_check_all2:", nrow(df_check_all2), "\n")  

```

## 7c. Create data  

### Create data from duplicated series   
- Variable names the same as in chunk 'Medians' in script 04  
```{r}

df_data_from_dupl <- df_check_all2 %>% 
  mutate(
    Dataset_pick = case_when(
      !is.na(val2_as) ~ "ICES",
      is.na(val1_as) ~ as.character(NA),
      val1_as > 0 ~ "EMODnet"),
    Value = case_when(
      Dataset_pick %in% "ICES" ~ val2_as,
      Dataset_pick %in% "EMODnet" ~ val1_as),
    Flag = case_when(
      Dataset_pick %in% "ICES" & u_loq2/n2 >= 0.5 ~ "<",
      Dataset_pick %in% "ICES" & u_loq2/n2 < 0.5 ~ as.character(NA),
      Dataset_pick %in% "EMODnet" & u_loq1/n1 >= 0.5 ~ "<",
      Dataset_pick %in% "EMODnet" & u_loq1/n1 < 0.5 ~ as.character(NA))
  ) %>%
  select(id, species, param, year, Dataset_pick, Value, Flag) %>%
  filter(!is.na(Dataset_pick))

cat("No of rows, df_data_from_dupl:", nrow(df_data_from_dupl), "\n")   

```


### Save data from duplicates   

- These are already medians and can be used directly in dat_04  

```{r}

if (FALSE){
  
  #
  # Check if the resulting data is identical to previous data
  #
  df_data_from_dupl_saved <- readRDS("Data/03_df_data_from_dupl.rds")
  nrow(df_data_from_dupl_saved)
  nrow(df_data_from_dupl)
  ncol(df_data_from_dupl_saved)
  ncol(df_data_from_dupl)
  
  for (i in 1:ncol(df_data_from_dupl)){
    identical(df_data_from_dupl[[i]], df_data_from_dupl_saved[[i]]) %>% print()
  }
  
}

if (TRUE){
  saveRDS(df_data_from_dupl, "Data/03_df_data_from_dupl_2025.rds")
}

```

## APPENDIX 1: Duplicate check, name missing  
- We would do this for Finland only   
- But it turnds out there are almost no Finnish data left 
```{r}

dat_a %>% filter(Country == "Finland") %>% pull(LatLong) %>% unique() %>% length()

#
# Only 4 Macoma balthica stations    
#
# Macoma balthica = Baltic clam or Baltic tellin, a small saltwater clam  
# Saduria entomon = benthic isopod crustacean
#
dat_a %>% 
  filter(Country == "Finland") %>% 
  distinct(SD_StationName, LatLong, WoRMS_scientificName, BASIS) %>%
  xtabs(~WoRMS_scientificName + BASIS, .)

# ...only very old data (until 1987) for this species     
dat_a %>% 
  filter(Country == "Finland" & WoRMS_scientificName == "Macoma balthica") %>% 
  distinct(Country, LatLong, WoRMS_scientificName, BASIS, Dataset, Year) %>%
  xtabs(~Year + BASIS + Dataset, .)

# One station in Lithuania with newer ICES data  
# Wet weight, dry weight fo rthe last years    
df_macoma <- dat_a %>% 
  filter(WoRMS_scientificName == "Macoma balthica" & Year >= 2010)
df_macoma %>%
  mutate(has_drywt = !is.na(DRYWT)) %>%
  distinct(Country, LatLong, WoRMS_scientificName, BASIS, Dataset, Year, has_drywt) %>%
  xtabs(~Year + has_drywt + BASIS + Dataset + Country, .)
mean(df_macoma$DRYWT, na.rm = TRUE)
# 19.96 %

if (FALSE){
  dat_a %>% 
    distinct(Country, LatLong, WoRMS_scientificName, BASIS, Dataset, Year) %>%
    filter(WoRMS_scientificName == "Macoma balthica" & Year >= 2000) %>%
    xtabs(~Year + BASIS + Dataset, .)
}



```

```{r}

if (FALSE){

df1_a <- dat_7 %>%
  filter(Country == "Finland") %>%
  mutate(Country = paste0(Country, "_", substr(Dataset, 1, 4))) %>%
  distinct(LatLong, Latitude, Longitude, Country, Dataset, Year)

# Unique for each position/dataset (years combined)
df2_a <- df1_a %>%
  distinct(LatLong, Latitude, Longitude, Country, Dataset) %>%
  arrange(LatLong, Latitude, Longitude)

df2_a_dist <- earth.dist.pairwise(df2_a, var_id = "LatLong") %>%
  arrange(dist) %>%
  # Add Dataset1 and Dataset2 columns
  left_join(df2_a %>% select(LatLong, Dataset) %>% rename(Dataset1 = Dataset),
            by = c("id1" = "LatLong")) %>%
  left_join(df2_a %>% select(LatLong, Dataset) %>% rename(Dataset2 = Dataset),
            by = c("id2" = "LatLong"))  

dist_limit <- 1
message("Distances < ", dist_limit, " km: ", 
        mean(df2_a_dist$dist < dist_limit, na.rm = TRUE)*100, " percent")

}


```

## APPENDIX 2: More checking of df_check_all2 (no changes)  

** This part was cut and saved separately as script 03extra  

## APPENDIX 3: Check positions for station with deviant MSFD region in one year   

- FRB65 (DK) has been registered with MSFD_region 'ATL' in 2006 and with 'ANS' in all other years (see median file 'Data/04_dat_medians.rds' produced by script 04)   
- This splits the time series, so 2006 data end up as one-year time series for each contaminant, and the other time series lack 2006  
- We remove these one-year time series in script 07 (the start of part 8, line no 345)    
- The code below shows that also the coordinates of this year shows that the coordinates are unusual from most years, so we just leave the fix in script 07 as it is    
- In other words, the 2006 data have been removed from this station, as identification was uncertain (due to position and given MSFD region)   

```{r}

# dat_7 <- readRDS("Data/03_dat_7_2025.rds") ???


dat_7 %>%
  filter(id  == "FRB65") %>%
  distinct(id, Longitude, Latitude, MYEAR, MSFD_region) %>%
  count(id, Longitude, Latitude, MSFD_region) %>%
  ggplot(aes(Longitude, Latitude)) +
  geom_text(aes(label = n, colour = MSFD_region))

```

