---
title: "Calculate medians for regression"
author: "DHJ"
date: "24 10 2021"
output: 
  html_document:
    keep_md: true  
    toc: true  
    toc_float: true
---


## 1. Packages    
```{r, message=FALSE, warning=FALSE, results='hide'}

library(dplyr)
library(purrr)
library(furrr)
library(tidyr)
library(ggplot2)
library(ggeasy)
library(lubridate)
library(forcats)

source("02_Fix_station_duplicates_functions.R")

```

## 2. Read data   

### Main data  
```{r}

dat_1_all <- readRDS("Data/03_dat_7.rds")

```

### Filtering   
- Removing EMODnet data for a couple of stations  
- See script 03 part 7b 'Check long distances'    
    -  MGAVSA" (Spain), 
    Mytilus edulis (ICES) + M. galloprovincialis (EMODnet)
    - "KBKLYN" (Denmark)
    Mytilus edulis 
```{r}

dat_1 <- dat_1_all %>%
  filter(!(id %in% c("MGAVSA", "KBKLYN") & Dataset %in% "EMODnet"))

nrow(dat_1_all)
nrow(dat_1)

```


### Duplication data set    
- Created in script 03  
```{r}

dat_medians_from_duplicates <- readRDS("Data/03_df_data_from_dupl.rds")

# df_duplicates <- readRDS(file = "Data/03_df_check_all.rds")

```

### Species-specific fat and dry weights  
- From OSPAR (https://ocean.ices.dk/ohat/trDocuments/2021/help_ac_basis_conversion.html )
```{r}

df_bivalve_dryweight <- read.csv(textConnection("
WoRMS_scientificName, Common_name, FATWT_species, DRYWT_species
Cerastoderma edule,	common cockle, NA, 19.0 		
Mya arenaria, softshell clam, 0.7, 14.8 		
Ruditapes philippinarum, manila clam, NA, 16.0 		
Mytilus edulis, blue mussel, 1.4, 16.3 		
Mytilus galloprovincialis, Mediteranean mussel, 2.2, 19.0 		
Crassostrea gigas, Pacific oyster, 2.1, 18.0 		
Ostrea edulis, native oyster, 1.8, 20.4 	
"), stringsAsFactors = FALSE)

# Check
tab <- table(dat_1$WoRMS_scientificName)
cat(100*mean(df_bivalve_dryweight$WoRMS_scientificName %in% rownames(tab)), 
    "% of names found in dat_1 \n")

dat_1 %>%
  filter(WoRMS_scientificName %in% df_bivalve_dryweight$WoRMS_scientificName) %>%
  count(WoRMS_scientificName)

tab[order(tab)] %>% rev() %>% head(30)

```
## 3. Medians for series that are only ICES or EMODnet       

### Check dataset for stations       
Table shows number of years per station and dataset  
```{r, results='hold'}

table(dat_1$Dataset)

tab_a <- dat_1 %>%
  mutate(Country = paste0(Country, "_", substr(Dataset, 1, 4))) %>%
  distinct(id, Country, Dataset, Year) %>% # View()
  
  xtabs(~id + Dataset, .)  

cat("Dataset with EMODnet data a \n\n")

sel <- tab_a[,"EMODnet"] > 0 & tab_a[,"ICES"] == 0
cat(sum(sel), "stations have data in EMODnet only (",  
  100*round(mean(sel),3), "% )\n\n")  
stations_emodnet <- rownames(tab_a[sel,])

sel <- tab_a[,"EMODnet"] == 0 & tab_a[,"ICES"] > 0
cat(sum(sel), "stations have data in ICES only (",  
  100*round(mean(sel),3), "% )\n\n")  
stations_ices <- rownames(tab_a[sel,])

# "MGAVSA"	Mytilus edulis
# "KBKLYN"	Mytilus edulis

sel <- tab_a[,"EMODnet"] > 0 & tab_a[,"ICES"] > 0
cat(sum(sel), "stations have data in both datasets (",
  100*round(mean(sel),3), "% )\n\n")  
stations_dupl <- rownames(tab_a[sel,])

```

### Basis   

* Note: also sets 'Value_as' = assessment value (dry weight for metals, otherwise wet weight)

```{r}

metals <- c("ZN", "CU", "CD", "HG", "CR", "NI", "PB", "AS")

dat_2 <- dat_1 %>%
  
  mutate(
      Value_dw = case_when(
        BASIS %in% "W" & DRYWT_MUNIT %in% "%" ~ Value/(DRYWT*0.01),
        BASIS %in% "L" & DRYWT_MUNIT %in% "%" ~ (Value*FATWT_species*0.01)/(DRYWT*0.01),
        BASIS %in% "D" ~ Value),
      Value_ww = case_when(
        BASIS %in% "W" ~ Value,
        BASIS %in% "L" ~ Value*(FATWT_species*0.01),
        BASIS %in% "D" & DRYWT_MUNIT %in% "%" ~ Value*(DRYWT*0.01)),
      # Assessment value  
      Value_as = case_when(
        PARAM %in% metals ~ Value_dw,
        !PARAM %in% metals ~ Value_ww)
      )

# dat_2 %>% filter(is.na(Value_dw)) %>% View
xtabs(~is.na(Value_dw), dat_2)
xtabs(~is.na(Value_ww), dat_2)
xtabs(~is.na(Value_as), dat_2)


if (FALSE)
  dat_2 %>% 
  select(Country, SD_StationCode, SD_StationName, MYEAR, PARAM, PARGROUP, MATRX, Value_orig, DRYWT, DRYWT_MUNIT, Value_dw, Value_ww, Value_as) %>% View("dat_2")

# Calculate dry and wet-weight basis (but DRYWT often lacking in one of the pair)

```

### Add LOQ value   
- per country/parameter/year (for ALL obs.)  
```{r}

dat_2_loqvalues <- dat_2 %>%
  group_by(Country, PARAM, Year) %>%
  summarise(
    n_below_loq = sum(!is.na(QFLAG)),
    LOQ_value_dw = case_when(
      n_below_loq > 0 ~ median(Value_dw[!is.na(QFLAG)]),         
      n_below_loq == 0 ~ as.double(NA)),
    LOQ_value_ww = case_when(
      n_below_loq > 0 ~ median(Value_ww[!is.na(QFLAG)]),         
      n_below_loq == 0 ~ as.double(NA)),
    .groups = "drop"
  ) %>%
  mutate(
    LOQ_value_as = case_when(
      PARAM %in% metals ~ LOQ_value_dw,
      !PARAM %in% metals ~ LOQ_value_ww)
  ) %>%
  select(-n_below_loq)


```

### Medians   

* Note: 'Value' of median data is the 'Value_as' = assessment value (dry weight for metals)

```{r}

dat_medians_nodupl <- dat_2 %>%
  filter(id %in% c(stations_emodnet, stations_ices)) %>%
  group_by(
    Country, MSFD_region, id, Dataset, WoRMS_scientificName, PARAM, Year, UNIT) %>%
  summarise(
    Value = median(Value_as),
    n = n(),
    Below_LOQ = sum(!is.na(QFLAG)),
    Flag = ifelse(Below_LOQ/n > 0.5, "<", as.character(NA)),
    .groups = "drop"
  ) 

```

#### Check/demonstrate for HG data  
```{r, results = 'hold'}
cat("----------------------------------------------------------------------------  \n")
cat("dat_1, ww basis compared with 20, dw basis compared with 129   \n")
cat("----------------------------------------------------------------------------  \n")

dat_1 %>%
  filter(MYEAR >= 2010,
         PARAM == "HG", 
         Country %in% c("Germany", "Denmark", "Sweden", "Poland"),
         BASIS == "W") %>%
  xtabs(~(Value > 20) + Country + BASIS, .)

dat_1 %>%
  filter(MYEAR >= 2010,
         PARAM == "HG", 
         Country %in% c("Germany", "Denmark", "Sweden", "Poland"),
         BASIS == "D") %>%
  xtabs(~(Value > 129) + Country + BASIS, .)

cat("----------------------------------------------------------------------------  \n")
cat("dat_2, ww basis compared with 20, dw basis compared with 129   \n")
cat("----------------------------------------------------------------------------  \n")

dat_2 %>%
  filter(MYEAR >= 2010,
         PARAM == "HG", 
         Country %in% c("Germany", "Denmark", "Sweden", "Poland")) %>%
  xtabs(~(Value_ww > 20) + Country, .)

dat_2 %>%
  filter(MYEAR >= 2010,
         PARAM == "HG", 
         Country %in% c("Germany", "Denmark", "Sweden", "Poland")) %>%
  xtabs(~(Value_dw > 129) + Country, .)

cat("----------------------------------------------------------------------------  \n")
cat("dat_medians_nodupl, Value (= dw basis for HG) compared with 129   \n")
cat("----------------------------------------------------------------------------  \n")

dat_medians_nodupl %>%
  filter(Year >= 2010,
         PARAM == "HG", 
         Country %in% c("Germany", "Denmark", "Sweden", "Poland")) %>%
  xtabs(~(Value > 129) + Country, .)


```


## 4. Combine with medians from series with both ICES and EMODnet

### Add medians 

```{r}

dat_country <- dat_1 %>%
  distinct(id, Country, MSFD_region)

dat_medians <- bind_rows(
  dat_medians_nodupl,
  dat_medians_from_duplicates %>%
    mutate(
      Dataset = "Both",
      UNIT = "ug/kg") %>%
    rename(
      WoRMS_scientificName = species, PARAM = param, Year = year) %>%
    left_join(dat_country, by = "id")
)

# check
# apply(!is.na(dat_medians_from_duplicates), 2, mean)
apply(!is.na(dat_medians), 2, mean)

```


### Add LOQ values  
```{r}

n1 <- nrow(dat_medians)
dat_medians <- dat_medians %>%
  left_join(
    dat_2_loqvalues,  by = c("Country", "PARAM", "Year")
  )
n2 <- nrow(dat_medians)

if (n1 != n2)
  stop("left_join changed row number")

```

## 5. Save  
```{r}

filename_medians <- "Data/04_dat_medians_2024.rds"

# Save backup
timestamp <- Sys.time() %>% gsub(" ", "_", .) %>% gsub(":", "-", .)
new_filename <- paste0(
  sub(".rds", "", filename_medians, fixed = TRUE),
  "_", timestamp, ".rds"
)
file.copy(filename_medians, new_filename)

# Save
saveRDS(dat_medians, filename_medians)

# old
# saveRDS(dat_medians_nodupl, "Data/05_dat_medians_nodupl")

```



## APPENDIX 

### Time series per country    
```{r}

brks <- c(1,10,25,100,1000,10000)

df_plot <- dat_medians %>%
  distinct(
    Country, id, Dataset, WoRMS_scientificName, Year, PARAM) %>%
  count(
    Country, Year) %>%
  mutate(
    n_cat = cut(n, breaks = brks)
  ) 

table(df_plot$n_cat)

ggplot(df_plot, aes(Year, Country, fill = n_cat)) +
  geom_tile() +
  scale_fill_viridis_d(option = "plasma") +
  labs(title = "")

```


### Checks  
```{r}

dat_medians_nodupl %>%
  distinct(
    Country, id, Dataset, WoRMS_scientificName, Year, PARAM) %>%
  count(
    Country, id, Dataset, WoRMS_scientificName, Year) %>%
  ggplot(aes(Year, id, fill = n)) +
  geom_tile()

```







