---
title: "Results from MAR001, 2021"
author: "DHJ"
date: "29 10 2021"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_float: yes
  word_document:
    toc: yes
editor_options: 
  chunk_output_type: console
---


```{r, echo = FALSE, message=FALSE, warning=FALSE, results='hide'}

library(dplyr)
library(purrr)
library(furrr)       # future_map
library(collateral)  # map_peacefully
library(tidyr)
library(ggplot2)
library(ggeasy)
library(ggdist)      # geom_dots (used in function 'plot_relclass_param2')
library(janitor)     # tabyl + adorn_...
library(knitr)       # kable
library(kableExtra)  # kable_classic
# library(lubridate)
library(forcats)
library(broom)       # tidy
library(R2jags)      # (in 'lm_leftcensored') 
library(glue)        # glue (in 'lm_leftcensored')  

# source("02_Fix_station_duplicates_functions.R")
# source("05_Time_trend_regression_functions.R")
source("06_MAR001_results_functions.R")

library(maps)   

knitr::opts_chunk$set(results = 'hold', echo = FALSE) 
# results = 'hold': collect the results from a chunk, print them together after the entire chunk  
# echo = FALSE: don't print code  


```


```{r data, echo=FALSE, results='hold'}

#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-
# Simple categorical trend + status (main data)
#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-

fn <- "Data/07_dat_status_trend.rds"
dat_status_trend <- readRDS(fn)

# Delete one PCB118 outlier
dat_status_trend <- dat_status_trend %>%
  filter(!(id %in% "56.02833333_20.83333333" & PARAM %in% "CB118"))
# nrow(dat_status_trend)

cat("File modified at: \n")
file.info(fn)$mtime


#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-
# Trend including slope, will be used for meta-analysis
#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-

dat_series_trend <- readRDS("Data/05_dat_series_trend.rds")

#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-
# Threshold values   
#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-

df_thresholds <- readxl::read_excel("Thresholds/MGR_05_df_thresholds.xlsx", sheet = "MGR")

if (FALSE){
  
  # for app
  file.copy("Data/05_dat_status_trend.rds", 
            "app01/05_dat_status_trend.rds", overwrite = TRUE)
  file.copy("Data/05_df_thresholds.rds", 
            "app01/05_df_thresholds.rds", overwrite = TRUE)
  file.copy("06_MAR001_results_2021_functions.R",
            "app01/06_MAR001_results_2021_functions.R", overwrite = TRUE)
  }

```

```{r}

# sequence in plots   
msfd_regions <- c("Black Sea",
                  "Adriatic Sea", "Ionia Sea and central Med.",
                  "Western Mediterranean Sea", 
                  "Greater North Sea", "Celtic Seas", 
                  "Bay of Biscay / Iberian coast",
                  "Baltic Sea", "Norwegian Sea", "Barents Sea", 
                  "Icelandic Ocean")


```


## Status assessment  

### Overall 
```{r, echo=FALSE, fig.width=10, fig.height=4}

dat_status_trend %>%
  mutate(PARAM = factor(
    PARAM, 
    levels = c("CD","CU","PB","HG","HCB","HCHG","DDEPP","BAP", "CB118"))
  ) %>%
  tabyl(Status, PARAM) %>%
  adorn_percentages("col") %>%
  adorn_pct_formatting(digits = 1) %>%
  adorn_ns() %>%
  kable() %>%
  kable_classic(full_width = F, html_font = "Arial Narrow")

```

## Status by region {.tabset}  


### North-East Atlantic Ocean
```{r, fig.width=10, fig.height=4}

table_status_region("North-East Atlantic Ocean")

```

### Baltic
```{r, fig.width=10, fig.height=4}

table_status_region("Baltic")

```


### Mediterranean  
```{r, fig.width=10, fig.height=4}

table_status_region("Mediterranean")

```

### Save  
```{r}

### SAVE ### 
# saveRDS(meta_status_region, "Data/06_meta_status_region.rds")
# saveRDS(meta_status_msfdregion, "Data/06_meta_status_msfdregion.rds")

```

## Status by MSFD region {.tabset}  


### Greater North Sea  
```{r}

table_status_msfdregion("Greater North Sea")

```

### Celtic Seas

```{r}

table_status_msfdregion("Celtic Seas")

```

### Bay of Biscay / Iberian coast  

```{r}

table_status_msfdregion("Bay of Biscay / Iberian coast")

```

### Baltic Sea

```{r}

table_status_msfdregion("Baltic Sea")

```

### Western Mediterranean Sea

```{r}

table_status_msfdregion("Western Mediterranean Sea")

```

### Adriatic Sea 

```{r}

table_status_msfdregion("Adriatic Sea")

```

### Ionia Sea and central Med. 

```{r}

table_status_msfdregion("Ionia Sea and central Med.")

```

### Norwegian 

```{r}

table_status_msfdregion("Norwegian Sea")

```



## Relative class   

### Calculation  
- See script 21 for details and testing   
```{r}

# Get data set with one line per PARAM, Group, and variable 'Basis_comb'
# Basis_comb = W-W, D-W, D-L etc. depening on basis of the two thresholds
dat_param_group <- dat_status_trend %>%
  count(PARAM, Group) %>%
  left_join(
    df_thresholds %>%
      mutate(Basis_comb = paste0(Thresh1_basis, "-", Thresh2_basis)) %>%
      select(PARAM, Group, Basis_comb)
    
  )


#
# Add 'Relclass' to data using function 'get_data_with_relclass'  
#
dat_status_trend_relclass <- pmap_dfr(
  list(
    dat_param_group$PARAM,
    dat_param_group$Group,
    dat_param_group$Basis_comb),
  get_data_with_relclass, 
  dataset = dat_status_trend) %>%
  mutate(PARAM = factor(PARAM, 
                        levels = c("CD","CU","PB","HG","CB118","HCB","HCHG","DDEPP","BAP"))
  ) %>%
  filter(!is.na(Relclass)) %>%
  rename(MSFD_region2 = MSFD_region) %>%
  # Add Overall trend
  mutate(
    MSFD_region = case_when(
      MSFD_region2 %in% "MWE" ~ "Western Mediterranean Sea",
      MSFD_region2 %in% "MAD" ~ "Adriatic Sea",
      MSFD_region2 %in% "MIC" ~ "Ionia Sea and central Med.",
      MSFD_region2 %in% "BLK" ~ "Black Sea",
      MSFD_region2 %in% "ABI" ~ "Bay of Biscay / Iberian coast",
      MSFD_region2 %in% "ACS" ~ "Celtic Seas",
      MSFD_region2 %in% "ANS" ~ "Greater North Sea",
      MSFD_region2 %in% "BAL" ~ "Baltic Sea",
      MSFD_region2 %in% "NOR" ~ "Norwegian Sea",
      MSFD_region2 %in% "BAR" ~ "Barents Sea",
      MSFD_region2 %in% "ICE" ~ "Icelandic Ocean")
  ) %>%
    mutate(MSFD_region = factor(MSFD_region, levels = msfd_regions))




  
```

### Plots  

```{r, fig.width=8,, fig.height=5}


ggplot(dat_status_trend_relclass, aes(x = PARAM, y = Relclass)) +
  geom_boxplot() +
  geom_jitter(aes(fill = Status), alpha = 0.3, width = 0.25, shape = 21) +
  scale_fill_manual(values = c(`1` = "green3", `2` = "yellow2", `3` = "red")) +
  theme_bw()

above_plot_border <- dat_status_trend_relclass %>%
  filter(Relclass > 10) %>%
  count(PARAM) %>%
  mutate(Relclass = 10.1, labeltext = paste("> 10:\nn =", n))

ggplot(dat_status_trend_relclass %>% filter(Relclass <= 10), 
       aes(x = PARAM, y = Relclass)) +
  geom_boxplot() +
  geom_jitter(aes(fill = Status), alpha = 0.3, width = 0.25, shape = 21) +
  scale_fill_manual(values = c(`1` = "green3", `2` = "yellow2", `3` = "red")) +
  scale_y_continuous(breaks = 0:10, minor_breaks = NULL, limits = c(0,10.8)) + 
  geom_text(data = above_plot_border, aes(label = labeltext), 
            vjust = -0.3, size = rel(3.5), lineheight = 0.9) +
  theme_bw() 



```


## Relative class by region {.tabset}   

### North-East Atlantic Ocean  
```{r}

plot_relclass_region("North-East Atlantic Ocean")

```

### Baltic

```{r}
plot_relclass_region("Baltic")
```

### Mediterranean

```{r}
plot_relclass_msfdregion("Mediterranean")
```


## Relative class by MSFD region {.tabset}   


### Greater North Sea  
```{r}

plot_relclass_msfdregion("Greater North Sea")

```



### Celtic Seas

```{r}

plot_relclass_msfdregion("Celtic Seas")

```

### Bay of Biscay / Iberian coast  

```{r}

plot_relclass_msfdregion("Bay of Biscay / Iberian coast")

```

### Baltic Sea

```{r}

plot_relclass_msfdregion("Baltic Sea")

```

### Western Mediterranean Sea

```{r}

plot_relclass_msfdregion("Western Mediterranean Sea")

```

### Adriatic Sea 

```{r}

plot_relclass_msfdregion("Adriatic Sea")

```

### Ionia Sea and central Med. 

```{r}

plot_relclass_msfdregion("Ionia Sea and central Med.")

```

### Norwegian 

```{r}

plot_relclass_msfdregion("Norwegian Sea")

```


## Relative class by parameter, regions {.tabset}    

### CB118  
```{r, fig.width=9, fig.height=6}

plot_relclass_param("CB118")

```

### HG  
```{r, fig.width=9, fig.height=6}

plot_relclass_param("HG")

```

### HCHG (Lindane)  

```{r, fig.width=9, fig.height=6}

plot_relclass_param("HCHG")

```


### DDEPP (indicator of DDT)  

```{r, fig.width=9, fig.height=6}

set.seed(43)
plot_relclass_param("DDEPP")

```



## Relative class by parameter, MSFD regions {.tabset}    



### CB118  
```{r, fig.width=9, fig.height=6}

# Set status to NA if the measurement is under LOQ (unless it's also Status = 1)
sel <- with(dat_status_trend_relclass, Below_LOQ & Status  %in% c("2", "3"))
mean(sel, na.rm = TRUE)
dat_status_trend_relclass$Status[sel] <- as.character(NA)

set.seed(43)

# Straight plot
plot_relclass_param("CB118", msfd_regions = TRUE)

# Custom plot   
gg <- plot_relclass_param("CB118", msfd_regions = TRUE)

# Custom plot
# Note: using plot_relclass_param2() instead of plot_relclass_param()  
# - dotplot instead of jittered scatter plot
gg <- plot_relclass_param2("CB118", msfd_regions = TRUE,
                           binwidth = 0.15, overflow = "compress")

gg2 <- gg +
  scale_y_continuous(
    limits = c(1,10.8),
    breaks = seq(0,10), 
    labels = c("", "0", "BAC", "EQS", paste0(2:8, "x EQS"))
  ) +
  theme(panel.grid.minor.y = element_blank()) +
  geom_vline(aes(xintercept = 3.5)) +
  geom_vline(aes(xintercept = 6.5))
gg2

if (TRUE){

  # Old version is named Relative_class_CB118_by_MSFDregion2.png", gg2, width = 9.5, height = 5, dpi = 400)
  ggsave("Figures/Relative_class_CB118_by_MSFDregion2.png", gg2, width = 9.5, height = 5, dpi = 400)
  
}


```


#### Data for excel  
```{r}


if (TRUE){

  #
  # Save plot data for EEA (start - writing excel is further down)
  #
  # boxplot data
  gg_boxplot <- plot_relclass_param2("CB118", msfd_regions = TRUE, get_boxplot_only = TRUE)
  dd_boxplot <- layer_data(gg_boxplot)
  
  x_labels <- c("Adriatic Sea", "Ionia Sea and central Med.", "Western Mediterranean Sea", 
                "Greater North Sea", "Celtic Seas", "Bay of Biscay / Iberian coast", 
                "Baltic Sea")
  
  plotdata_boxplot <- dd_boxplot %>%
    select(x, ymin, lower, middle, upper, ymax) %>%
    rename(
      `lower end of bar`= ymin,
      `lower end of box`= lower,
      `line inside box`= middle,
      `upper end of box`= upper,
      `upper end of bar`= ymax) %>%
    mutate(
      x = as.numeric(x),
      x_label = x_labels[x], .before = 1)
  
  # test <- layer_data(gg2, 2)
  # nrow(test)
  # ggplot(test, aes(x, y, shape = factor(shape))) +
  #   geom_point()
  
  # point colours
  plotdata_colors <- data.frame(
    Status = 1:3,
    colour = c("green", "yellow", "red")
  )
  
  # point data
  gg2 <- plot_relclass_param("CB118", msfd_regions = TRUE, get_points_only = TRUE)
  plotdata_points <- dat_status_trend_relclass %>%
    filter(PARAM %in% "CB118") %>%
    select(MSFD_region, Relclass, Status) %>%
    mutate(
      x = as.numeric(MSFD_region),
      Status = as.numeric(Status)) %>%
    left_join(plotdata_colors) %>%
    mutate(
      colour = case_when(
        is.na(colour) ~ "grey",
        TRUE ~ colour),
      shape = case_when(
        Below_LOQ ~ "triangle pointing down",
        TRUE ~ "circle")
    ) %>%
    rename(x_label = MSFD_region, 
           y = Relclass) %>%
    select(x, x_label, y, colour)
  
  plotdata_yscale <- data.frame(
    y = 0:9,
    y_label = c("0", "BAC", "EQS", paste0(2:8, "x EQS"))
  )

  # Save to excel
  # Manual changes: change column with, reduce number of digits
  writexl::write_xlsx(
    list(points = plotdata_points, 
         boxplot = plotdata_boxplot, `y scale` = plotdata_yscale),
    "Figures/2023/PCB118_by_MSFDregion_plotdata.xlsx" 
  )

  if (FALSE){
    
    # test box plot
    ggplot(plotdata_boxplot, aes(x, group = x)) +
      geom_boxplot(aes(lower=lower, upper=upper, middle=middle, ymin=ymin, ymax=ymax), stat = "identity")
    ggplot(plotdata_boxplot %>% mutate(MSFD_region = fct_inorder(MSFD_region)), 
           aes(MSFD_region, group = MSFD_region)) +
      geom_boxplot(aes(lower=lower, upper=upper, middle=middle, ymin=ymin, ymax=ymax), stat = "identity")
    
    # test points
    ggplot() +
      geom_jitter(data = plotdata_points, aes(x,y, color = colour))

    # test box plot + points
    ggplot(plotdata_boxplot, aes(x)) +
      geom_boxplot(aes(group = x, lower=lower, upper=upper, middle=middle, ymin=ymin, ymax=ymax), 
                   stat = "identity") +
      geom_point(data = plotdata_points, aes(x, y, color = colour))
    
  }
  ggplot2::ggplot_build(gg2)
  
  y_axis_labels <- data.frame(
    y = seq(0,10), 
    labels = c("", "0", "BAC", "EQS", paste0(2:8, "x EQS"))
  )


}


```


### HG  
```{r, fig.width=9, fig.height=6}

set.seed(43)
plot_relclass_param("HG", msfd_regions = TRUE)

if (TRUE){

  # Custom plot   

  
  set.seed(43)
  gg <- plot_relclass_param("HG", msfd_regions = TRUE)

  gg2 <- gg +
    scale_y_continuous(
      limits = c(1,10.8),
      breaks = seq(0,10), 
      labels = c("", "0", "EQS", "MPC", paste0(2:8, "x EAC"))
    ) +
    geom_vline(aes(xintercept = 1.5)) +
    geom_vline(aes(xintercept = 4.5)) +
    geom_vline(aes(xintercept = 7.5))

  ggsave("Figures/Relative_class_HG_by_MSFDregion.png", gg2, width = 8, height = 5, dpi = 400)
  
}

```

### HCHG (Lindane)  

```{r, fig.width=9, fig.height=6}

set.seed(43)
plot_relclass_param("HCHG", msfd_regions = TRUE)

```



### DDEPP (indicator of DDT)  

```{r, fig.width=9, fig.height=6}

set.seed(43)
plot_relclass_param("DDEPP", msfd_regions = TRUE)

```


## Trends table  

```{r}

dat_status_trend %>%
  mutate(PARAM = factor(
    PARAM, 
    levels = c("CD","CU","PB","HG","HCB","HCHG","DDEPP","BAP", "CB118"))
  ) %>%
  tabyl(Trend) %>%
  adorn_percentages("col") %>%
  adorn_pct_formatting(digits = 1) %>%
  adorn_ns() %>%
  kable() %>%       # add region as caption  
  kable_classic(full_width = F, html_font = "Arial Narrow") 

```


### Overall 
```{r, echo=FALSE}

dat_status_trend %>%
  mutate(PARAM = factor(
    PARAM, 
    levels = c("CD","CU","PB","HG","HCB","HCHG","DDEPP","BAP", "CB118"))
  ) %>%
  tabyl(Trend, PARAM) %>%
  adorn_percentages("col") %>%
  adorn_pct_formatting(digits = 1) %>%
  adorn_ns() %>%
  kable() %>%       # add region as caption  
  kable_classic(full_width = F, html_font = "Arial Narrow") 

```
#### Save as Excel 

* Percentages plus total numbers  
* Used in assessment text  

```{r}

tab1 <- dat_status_trend %>%
  mutate(PARAM = factor(
    PARAM, 
    levels = c("CD","CU","PB","HG","HCB","HCHG","DDEPP","BAP", "CB118"))
  ) %>%
  tabyl(Trend, PARAM)

# Percentages  
tab2a <- tab1 %>%
  adorn_percentages("col") %>%
  adorn_pct_formatting(digits = 1, affix_sign = FALSE)

# Numbers (only to get the totals on row 5)  
tab2b <- tab1 %>%
  adorn_totals()
# Turn numbers to characters
for (col in 1:ncol(tab2b))
  tab2b[,col] <- as.character(tab2b[,col])

# Combine
tab_report <- bind_rows(tab2a,  tab2b[5,])

writexl::write_xlsx(tab_report, "Figures/Trends_overall.xlsx")

tab_report

```


## Trend tables by region  


### North-East Atlantic Ocean
```{r, fig.width=10, fig.height=4}

table_trends_region("North-East Atlantic Ocean")

```

### Baltic
```{r, fig.width=10, fig.height=4}

table_trends_region("Baltic")

```


### Mediterranean  
```{r, fig.width=10, fig.height=4}

table_trends_region("Mediterranean")

```




## Trend tables by MSFD region{.tabset} 


### Greater North Sea  
```{r}

table_trends_msfdregion("Greater North Sea")

```

### Celtic Seas

```{r}

# debugonce(table_trends_msfdregion)
table_trends_msfdregion("Celtic Seas")

```

### Bay of Biscay / Iberian coast  

```{r}

table_trends_msfdregion("Bay of Biscay / Iberian coast")

```

### Baltic Sea

```{r}

table_trends_msfdregion("Baltic Sea")

```

### Western Mediterranean Sea

```{r}

table_trends_msfdregion("Western Mediterranean Sea")

```

### Adriatic Sea 

```{r}

table_trends_msfdregion("Adriatic Sea")

```

### Ionia Sea and central Med. 

```{r}

table_trends_msfdregion("Ionia Sea and central Med.")

```

### Norwegian Sea  

```{r}

table_trends_msfdregion("Norwegian Sea")

```




```{r}

#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#ov#o#o#o
#
# Trends by region   
#  - Meta-analysis of trends by station  
#
#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#o#ov#o#o#o


```


```{r, echo=FALSE, fig.width=8, fig.height=3}

#
# Perform meta-analysis
#

meta_trends_overall <- dat_series_trend %>%
  filter(!is.na(Slope)) %>%
  mutate(Dummy = 1) %>%
  nest_by(PARAM) %>% # str(2)
  mutate(meta_result = list(get_slope_estimates_grouped(data, "Dummy"))) %>%
  unnest(meta_result) %>%
  mutate(
    Change_perc = 100*(10^(Estimate)-1),
    Change_perc_10yr = 100*(10^(10*Estimate)-1),
    Change_perc_10yr_lo = 100*(10^(10*CI_lower)-1),
    Change_perc_10yr_up = 100*(10^(10*CI_upper)-1)
  ) %>%
  # Add Overall trend
  mutate(
    `Overall trend` = case_when(
      CI_upper < 0 ~ "Downward",
      CI_lower > 0 ~ "Upward",
      TRUE ~ "No sign. pattern")
  ) 

meta_trends_region <- dat_series_trend %>%
  filter(!is.na(Slope)) %>%
  nest_by(PARAM) %>% # str(2)
  mutate(meta_result = list(get_slope_estimates_grouped(data, "Region"))) %>%
  unnest(meta_result) %>%
  mutate(
    Change_perc = 100*(10^(Estimate)-1),
    Change_perc_10yr = 100*(10^(10*Estimate)-1),
    Change_perc_10yr_lo = 100*(10^(10*CI_lower)-1),
    Change_perc_10yr_up = 100*(10^(10*CI_upper)-1)
  ) %>%
  # Add Overall trend
  mutate(
    `Overall trend` = case_when(
      CI_upper < 0 ~ "Downward",
      CI_lower > 0 ~ "Upward",
      TRUE ~ "No sign. pattern")
  ) 

meta_trends_msfdregion <- dat_series_trend %>%
  filter(!is.na(Slope)) %>%
  nest_by(PARAM) %>% # str(2)
  mutate(meta_result = list(get_slope_estimates_grouped(data, "MSFD_region"))) %>%
  unnest(meta_result) %>%
  mutate(
    Change_perc = 100*(10^(Estimate)-1),
    Change_perc_10yr = 100*(10^(10*Estimate)-1),
    Change_perc_10yr_lo = 100*(10^(10*CI_lower)-1),
    Change_perc_10yr_up = 100*(10^(10*CI_upper)-1)
  ) %>%
  rename(MSFD_region2 = MSFD_region) %>%
  # Add Overall trend
  mutate(
    `Overall trend` = case_when(
      CI_upper < 0 ~ "Downward",
      CI_lower > 0 ~ "Upward",
      TRUE ~ "No sign. pattern"),
    MSFD_region = case_when(
      MSFD_region2 %in% "MWE" ~ "Western Mediterranean Sea",
      MSFD_region2 %in% "MAD" ~ "Adriatic Sea",
      MSFD_region2 %in% "MIC" ~ "Ionia Sea and central Med.",
      MSFD_region2 %in% "BLK" ~ "Black Sea",
      MSFD_region2 %in% "ABI" ~ "Bay of Biscay / Iberian coast",
      MSFD_region2 %in% "ACS" ~ "Celtic Seas",
      MSFD_region2 %in% "ANS" ~ "Greater North Sea",
      MSFD_region2 %in% "BAL" ~ "Baltic Sea",
      MSFD_region2 %in% "NOR" ~ "Norwegian Sea",
      MSFD_region2 %in% "BAR" ~ "Barents Sea",
      MSFD_region2 %in% "ICE" ~ "Icelandic Ocean")
  ) 

col_scale <- c("Downward"  = "green3", "Upward" = "red", "No sign. pattern" = "black")



xtabs(~ addNA(MSFD_region) + MSFD_region2, meta_trends_msfdregion)

```


```{r}

### SAVE ### 

saveRDS(meta_trends_region, "Data/06_meta_trends_region.rds")
saveRDS(meta_trends_msfdregion, "Data/06_meta_trends_msfdregion.rds")

```


## Trend meta-analysis by region {.tabset}


### CD  
```{r, fig.width=6, fig.height=2}

plot_slope_region("CD")

```

### CU  
```{r, fig.width=6, fig.height=2}

plot_slope_region("CU")

```

### PB  
```{r, fig.width=6, fig.height=2}

plot_slope_region("PB")

```

### HG  
```{r, fig.width=6, fig.height=2}

plot_slope_region("HG")

```

### HCB  
```{r, fig.width=6, fig.height=2}

plot_slope_region("HCB")

```

### HCHG (Lindane)  
```{r, fig.width=6, fig.height=2}

plot_slope_region("HCHG")

```

### DDEPP (indicator of DDT)  
```{r, fig.width=6, fig.height=2}

plot_slope_region("DDEPP")

```

### CB118  
```{r, fig.width=6, fig.height=2}

plot_slope_region("CB118")

```

### BAP (Benzo[a]pyrene)  
```{r, fig.width=6, fig.height=2}

plot_slope_region("BAP")

```


## Trend meta-analysis by MSFD region {.tabset}


### CD  
```{r, fig.width=6, fig.height=3}

plot_slope_msfdregion("CD")

```

### CU  
```{r, fig.width=6, fig.height=3}

plot_slope_msfdregion("CU")

```

### PB  
```{r, fig.width=6, fig.height=3}

plot_slope_msfdregion("PB")

```

### HG  
```{r, fig.width=6, fig.height=3}

plot_slope_msfdregion("HG")  

if (TRUE){

  # Custom plot   

  
  set.seed(43)
  gg <- plot_slope_msfdregion("HG")  

  gg2 <- gg +
    scale_y_continuous(
      limits = c(1,10.8),
      breaks = seq(0,10), 
      labels = c("", "0", "EQS", "MPC", paste0(2:8, "x EAC"))
    ) +
    geom_vline(aes(xintercept = 1.5)) +
    geom_vline(aes(xintercept = 4.5)) +
    geom_vline(aes(xintercept = 7.5))

  ggsave("Figures/Meta_trend_HG_by_MSFDregion.png", gg, width = 6, height = 3, dpi = 400)
  
}



```

### HCB  
```{r, fig.width=6, fig.height=3}

plot_slope_msfdregion("HCB")

```

### HCHG (Lindane)  
```{r, fig.width=6, fig.height=3}

plot_slope_msfdregion("HCHG")

```

### DDEPP (indicator of DDT)  
```{r, fig.width=6, fig.height=3}

plot_slope_msfdregion("DDEPP")


if (TRUE){

  # Custom plot   

  
  set.seed(43)
  gg <- plot_slope_msfdregion("DDEPP")  

  ggsave("Figures/Meta_trend_DDEPP_by_MSFDregion.png", gg, width = 6, height = 3, dpi = 400)
  
}

```

### CB118  
```{r, fig.width=6, fig.height=3}

plot_slope_msfdregion("CB118")

```


#### Save plot
```{r, fig.width=6, fig.height=3}


if (TRUE){

  # Custom plot   

  set.seed(43)
  gg <- plot_slope_msfdregion("CB118")  

  ggsave("Figures/Meta_trend_CB118_by_MSFDregion.png", gg, width = 6, height = 3, dpi = 400)
  
}

```


#### Save as excel (plot data for this plot and boxplot)   
```{r, fig.width=6, fig.height=3}

if (TRUE){
  
  dd <- layer_data(gg) %>%
    arrange(desc(x)) %>%
    mutate(MSFD_region = c("Greater North Sea", "Celtic Seas", 
                           "Bay of Biscay / Iberian coast", "Baltic Sea"))
  
  #
  # Save plot data
  # 
  # NOTE: 'regression_lines' and 'labels' must already have been created (in memory) in 
  #   'Relative class by parameter, MSFD regions' - CB118
  #
  
  writexl::write_xlsx(
    list(
      points = plotdata_points,
      boxplot = plotdata_boxplot,
      `y axis labels` = y_axis_labels,
      `inset barplot` = dd
    ),
    "Figures/06_Relative_class_CB118_by_MSFDregion_plotdata.xlsx"
  )
  
  

}


```

### BAP (Benzo[a]pyrene)  
```{r, fig.width=6, fig.height=3}

plot_slope_region("BAP")

```

### Remake trends table with addition    

#### Extend the already made Trends table     

* Percentages plus total numbers  
* Used in assessment text  

```{r}

# Made higher up in this script
tab_report <- readxl::read_excel("Figures/Trends_overall.xlsx")

# Parameters

params <- c("CD","CU","PB","HG","HCB","HCHG","DDEPP","BAP", "CB118")

# Overall trends 
trend_overall_meta <- meta_trends_overall %>%
  select(PARAM, Dummy, Change_perc_10yr, Change_perc_10yr_lo, Change_perc_10yr_up, `Overall trend`) %>%
  filter(PARAM %in% params)

# Overall trends , reformat 1
df1 <- trend_overall_meta %>%
    mutate(
      PARAM = factor(
      PARAM, 
      levels = c("CD","CU","PB","HG","HCB","HCHG","DDEPP","BAP", "CB118")),
      Change_txt = paste0(
        round(Change_perc_10yr, 0), " (",
        round(Change_perc_10yr_lo, 0), ",", round(Change_perc_10yr_up, 0), ")")
    ) %>%
    select(PARAM, Change_txt,  `Overall trend`)

# Overall trends , reformat 2
df2 <- bind_rows(
  df1 %>% 
    select(PARAM, `Overall trend`) %>%
    pivot_wider(names_from = PARAM, values_from = `Overall trend`),
  
  df1 %>% 
    select(PARAM, Change_txt) %>%
    pivot_wider(names_from = PARAM, values_from = Change_txt)
)

# Overall trends , reformat 3
df3 <- data.frame(
  Trend = c("Overall trend", "Change/decade (%)"),
  df2)

# Combine
tab_report_extended <- bind_rows(tab_report, df3)

tab_report_extended

```
#### Save excle
```{r}


# Made higher up in this script
# tab_report <- readxl::read_excel("Figures/Trends_overall.xlsx")

```

## Maps {.tabset}

```{r status_colors}

# Colors used for status  
colors_status <- RColorBrewer::brewer.pal(8, "Set1")[c(2,6,5,1)] # blue, yellow, orange, red
colors_status[3] <- "orange"   # sharper orange
colors_status[4] <- "red"      # sharper red
names(colors_status) <- c("blue", "yellow", "orange", "red")

```

```{r background map}

# Background map  
my_map <- map_data("world")

```

```{r make_single_plot, echo=FALSE}  

# NOTE: At the end of this part, we also make plots to put together to a combined plot
# See "map_param_for_combined_plot" below (just a modified version of 'map_param')

```

### CD

```{r, echo=FALSE, fig.width = 9, fig.height = 7}

# debugonce(map_param)
map_param("CD")

# or:
# map_param("CD", "custom")

```


### CU

```{r, echo=FALSE, fig.width = 9, fig.height = 7}
map_param("CU")
```

### PB

```{r, echo=FALSE, fig.width = 9, fig.height = 7}
map_param("PB")
```

### HG

```{r, echo=FALSE, fig.width = 9, fig.height = 7}
map_param("HG")
```

### HCB

```{r, echo=FALSE, fig.width = 9, fig.height = 7}
map_param("HCB")
```

### HCHG (Lindane)

```{r, echo=FALSE, fig.width = 9, fig.height = 7}
map_param("HCHG")
```

### DDEPP (indicator of DDT)

```{r, echo=FALSE, fig.width = 9, fig.height = 7}
map_param("DDEPP")
```


### CB118

```{r, echo=FALSE, fig.width = 9, fig.height = 7}
map_param("CB118")
```


## Combined map  

### Test a single map    
```{r make_combined_plot1, echo=FALSE, fig.width = 7, fig.height = 7}

# map_param_for_combined_plot_s <- safely(map_param_for_combined_plot)

# Test to get x and y limits for CD plot (which covers the biggest area)
# debugonce(map_param_for_combined_plot)
test <- map_param_for_combined_plot("CD")
map_param_for_combined_plot("CD", legends = "shape")
map_param_for_combined_plot("CD", legends = "none")
# test <- map_param_for_combined_plot("CD", labeltype = "custom")
# test <- map_param_for_combined_plot("CU")

if (FALSE){
  # get x and y limits (used in code above)
  layer_scales(test)$x$range$range %>% dput()
  layer_scales(test)$y$range$range %>% dput()
}

test

```


### Combined map, label "low, moderate, high", ver. 1: colour legend in map  
```{r, fig.width=10, fig.height=10}

map_list <- 
  c("CD","CU","PB","HG","HCB","HCHG","DDEPP","BAP", "CB118") %>%
  purrr::map(map_param_for_combined_plot)

# Make shape legend  
shape_legend <- map_param_for_combined_plot("CD", legends = "shape") %>%
  cowplot::get_legend()

gg <- cowplot::plot_grid(
  cowplot::plot_grid(plotlist = map_list, nrow = 3),
  cowplot::plot_grid(shape_legend, NULL, NULL, nrow = 1),
  nrow = 2, rel_heights = c(1,0.2)
)

ggsave("Figures/06_Combined_map_ver1.png", gg, 
       width = 13, height = 13, 
       bg = "white")

gg

```


### Combined map, label "low, moderate, high", ver. 2: no legend in map  
```{r, fig.width=10, fig.height=10}

map_list <- 
  c("CD","CU","PB","HG","HCB","HCHG","DDEPP","BAP", "CB118") %>%
  purrr::map(map_param_for_combined_plot, legends = "none")

# Make fill legend  
fill_legend <- map_param_for_combined_plot("CD", legends = "fill") %>%
  cowplot::get_legend()

# Make shape legend  
shape_legend <- map_param_for_combined_plot("CD", legends = "shape") %>%
  cowplot::get_legend()

gg <- cowplot::plot_grid(
  cowplot::plot_grid(plotlist = map_list, nrow = 3),
  cowplot::plot_grid(fill_legend, shape_legend, NULL, nrow = 1),
  nrow = 2, rel_heights = c(1,0.2)
)

ggsave("Figures/06_Combined_map_ver2.png", gg, 
       width = 13, height = 13, 
       bg = "white")

gg

``` 


### Combined map, custom labels     
```{r, fig.width=10, fig.height=10}

map_list <- 
  c("CD","CU","PB","HG","HCB","HCHG","DDEPP","BAP", "CB118") %>%
  purrr::map(map_param_for_combined_plot, labeltype = "custom")

# Make legend
# not needed, reuse existing

gg <- cowplot::plot_grid(
  cowplot::plot_grid(plotlist = map_list, nrow = 3),
  cowplot::plot_grid(shape_legend, NULL, NULL, nrow = 1),
  nrow = 2, rel_heights = c(1,0.2)
)

ggsave("Figures/06_Combined_map_customlabels.png", gg, 
       width = 13, height = 13, 
       bg = "white")

ggsave("Figures/06_Combined_map_customlabels_res500.png", gg,
       width = 13, height = 13, dpi = 500,
       bg = "white")

gg

```


## Levels by species  

### Threshold 1  
```{r}

plot_level_vs_species("CD")
plot_level_vs_species("CU")
plot_level_vs_species("PB")
plot_level_vs_species("HG")
plot_level_vs_species("HCB")
plot_level_vs_species("HCHG")    
plot_level_vs_species("DDEPP")
plot_level_vs_species("CB118")    
# plot_level_vs_species("CB118", log = FALSE)  
plot_level_vs_species("BAP")

# param <- "CB118"
# dat_status_3 %>% filter(PARAM == param) %>% View(param)

```

### Threshold 2  
```{r}

plot_level_vs_species("CD", threshold = 2)
plot_level_vs_species("HG", threshold = 2)
plot_level_vs_species("HCHG", threshold = 2)  
plot_level_vs_species("CB118", threshold = 2)  

```



## Levels by region     

### Threshold 1  
```{r}

plot_level_vs_region("CD")
plot_level_vs_region("CU")
plot_level_vs_region("PB")
plot_level_vs_region("HG")
plot_level_vs_region("HCB")   
plot_level_vs_region("HCHG")    
plot_level_vs_region("DDEPP")
plot_level_vs_region("CB118")    
plot_level_vs_region("BAP")

# param <- "CB118"
# dat_status_3 %>% filter(PARAM == param) %>% View(param)

```


### Threshold 2  
```{r}

plot_level_vs_region("CD", threshold = 2)
plot_level_vs_region("HG", threshold = 2)
plot_level_vs_region("HCHG", threshold = 2)  
plot_level_vs_region("CB118", threshold = 2)  

```



## Levels by Country       

### Threshold 1  
```{r}

plot_level_vs_country("CD")
plot_level_vs_country("CU")
plot_level_vs_country("PB")
plot_level_vs_country("HG")
plot_level_vs_country("HCB")   
plot_level_vs_country("HCHG")    
plot_level_vs_country("DDEPP")
plot_level_vs_country("CB118")    
plot_level_vs_country("BAP")

# param <- "CB118"
# dat_status_3 %>% filter(PARAM == param) %>% View(param)

```


### Threshold 2  
```{r}

plot_level_vs_country("CD", threshold = 2)
plot_level_vs_country("HG", threshold = 2)
plot_level_vs_country("HCHG", threshold = 2)  
plot_level_vs_country("CB118", threshold = 2)  

```

