---
title: "07_Status_classification"
author: "DHJ"
date: "5 11 2021"
output: 
  html_document:
    toc: true
    toc-float: true  
  github_document:
    toc: true
---

**Classifies current levels according to status**  

* Also adds trend data to a common file   
* The criteria used for classification is that time series that did not have at least 3 years of data since 2010  
    - Note: for the first version of the assessment, this criteria was not implemented properly   
    - This was corrected on 2022-06-17 (commit 1426b98)   
    - For comparison, the old saved file from this script was saved as '07_dat_status_trend_BEFORE_UPDATE_2022-06-17.rds'  
    - For comparison, this is used in the result file '06_MAR001_results_2021_BEFORE_UPDATE_2022-06-17.Rmd'  



## 1. Packages    
```{r, message=FALSE, warning=FALSE, results='hide'}

library(dplyr)
library(purrr)
library(tidyr)
library(ggplot2)
library(ggeasy)
# library(lubridate)
library(forcats)


# If package maps is loaded, unload it (messes up purrr::map)
if ("maps" %in% (.packages()))
  detach("package:maps", unload=TRUE)


```


## 2. Data   
*  Thresholds are at least partly from  
https://dome.ices.dk/ohat/trDocuments/2021/help_ac_biota_contaminants.html   
```{r}

### Medians (main data)
fn <- "Data/04_dat_medians.rds"
dat_medians <- readRDS(fn) %>%
  filter(Value > 0) %>%                      # remove zeros (due to EMODnet data < LOQ)
  mutate(
    Log_value = log10(Value),         # log10-transformed values will be used
    Region = case_when(
      MSFD_region %in% "BAL" ~ "Baltic",
      MSFD_region %in% c("MWE","MAD","MIC") ~ "Mediterranean",
      MSFD_region %in% "BLK" ~ "Black Sea",
      TRUE ~ "North-East Atlantic Ocean")
    )
cat("Medians updated: \n")
print(file.info(fn)[1, "mtime"])

# Data (for dry weight)  
dat_ind <- readRDS("Data/03_dat_7.rds")

# regression results  
fn <- "Data/05_dat_series_trend.rds"
dat_series_trend <- readRDS(fn)
cat("Regressions updated: \n")
print(file.info(fn)[1, "mtime"])

### Threshold values   
df_thresholds <- readxl::read_excel("Thresholds/MGR_05_df_thresholds.xlsx", sheet = "MGR")


```

### Make metadata 'dat_series'     
```{r}

dat_coordinates <- dat_ind %>%
  group_by(id) %>%
  filter(!is.na(Longitude)) %>%
  summarise(
    across(c(Longitude, Latitude), median), .groups = "drop")

dat_series <- dat_medians %>%
  group_by(Country, MSFD_region, Region, id, WoRMS_scientificName, PARAM, Dataset) %>%
  summarise(
    first_year = min(Year),
    last_year = max(Year),
    n_years = length(unique(Year)), .groups = "drop"
  ) %>%
  left_join(
    dat_coordinates, by = "id"
  )

ggplot(dat_series, aes(last_year, n_years)) +
  geom_bin2d()

dat_series %>%
  xtabs(~(last_year > 2015) + (n_years >= 5), .)

```

## 3. Criteria  

```{r}

min_year <- 2010  
min_no_years_except_med <- 3  
min_no_years_mediterr <- 1  

metals <- c("ZN", "CU", "CD", "HG", "CR", "NI", "PB", "AS")
oysters <- c("Crassostrea gigas", "Ostrea edulis")


```


## 4. Calculate status  

### Create dat_status_1       

- Percentiles use R type 8, recommended by Rob Hyndman. Hyndman, R. J. and Fan, Y. (1996) Sample quantiles in statistical packages, American Statistician 50, 361–365. doi: 10.2307/2684934.  
```{r}

dat_drywt_med <- dat_ind %>%
  group_by(Country, id, WoRMS_scientificName, Year) %>%
  summarise(
    DRYWT = median(DRYWT, na.rm = TRUE),
    FATWT_species = median(FATWT_species, na.rm = TRUE),
    .groups = "drop")


dat_status_1 <- dat_medians %>%
  filter(Year >= min_year) %>%
  left_join(dat_drywt_med, by = c("Country", "id", "WoRMS_scientificName", "Year")) %>% # View 
  mutate(
    BASIS = case_when(
      PARAM %in% metals ~ "D",
      !PARAM %in% metals ~ "W"),
    Value = 10^Log_value,
    Value_ww = case_when(
      BASIS == "D" ~ Value*(0.01*DRYWT),
      BASIS == "W" ~ Value),
    Value_dw = case_when(
      BASIS == "D" ~ Value,
      BASIS == "W" ~ Value/(0.01*DRYWT)),
    Value_fw = case_when(
      BASIS == "D" ~ Value/(0.01*FATWT_species)*(0.01*DRYWT),
      BASIS == "W" ~ Value/(0.01*FATWT_species))
  ) %>%
  group_by(Country, MSFD_region, id, Dataset, WoRMS_scientificName, PARAM) %>%
  summarise(
    n_years = length(unique(Year)),
    median_value = median(Value, na.rm = TRUE), 
    median_value_ww = median(Value_ww, na.rm = TRUE), 
    median_value_dw = median(Value_dw, na.rm = TRUE), 
    median_value_fw = median(Value_fw, na.rm = TRUE), 
    perc90_value = quantile(Value, 0.9, type = 8, na.rm = TRUE), 
    perc90_value_ww = quantile(Value_ww, 0.9, type = 8, na.rm = TRUE), 
    perc90_value_dw = quantile(Value_dw, 0.9, type = 8, na.rm = TRUE), 
    perc90_value_fw = quantile(Value_fw, 0.9, type = 8, na.rm = TRUE), 
    .groups = "drop"
  ) %>%
  ungroup() %>%
  mutate(
    status_analysis = case_when(
      MSFD_region %in% c("MAD", "MAL", "MIC", "MWE") & n_years >= min_no_years_mediterr ~ "Yes",
      MSFD_region %in% c("MAD", "MAL", "MIC", "MWE") & n_years < min_no_years_mediterr ~ "No",
      n_years >= min_no_years_except_med ~ "Yes",
      n_years <= min_no_years_except_med ~ "No",
      TRUE ~ "No"
    )
  )

```



### dat_status_2
```{r}

df_bivalve_dryweight <- readRDS("Data/03_df_bivalve_dryweight.rds")

dat_status_2 <- dat_status_1 %>%
  mutate(
    BASIS = case_when(
      PARAM %in% metals ~ "D",
      !PARAM %in% metals ~ "W"
    ),
    Group = ifelse(WoRMS_scientificName %in% oysters, "Oyster", "Mussel")
  ) %>%
  left_join(
    df_thresholds, by = c("PARAM", "Group")) %>%
  left_join(
    df_bivalve_dryweight, by = c("WoRMS_scientificName")) %>% 
  mutate(
    Status = case_when(
      Thresh1_basis %in% "W" & perc90_value_ww < Thresh1 ~ 1,
      Thresh1_basis %in% "W" & perc90_value_ww >= Thresh1 & is.na(Thresh2_basis) ~ 2,
      Thresh1_basis %in% "W" & perc90_value_ww >= Thresh1 & Thresh2_basis %in% "W" & perc90_value_ww < Thresh2 ~ 2,
      Thresh1_basis %in% "W" & perc90_value_ww >= Thresh1 & Thresh2_basis %in% "W" & perc90_value_ww >= Thresh2 ~ 3,
      Thresh1_basis %in% "W" & perc90_value_ww >= Thresh1 & Thresh2_basis %in% "L" & perc90_value_fw < Thresh2 ~ 3,
      Thresh1_basis %in% "W" & perc90_value_ww >= Thresh1 & Thresh2_basis %in% "L" & perc90_value_fw >= Thresh2 ~ 3,
      Thresh1_basis %in% "D" & perc90_value_dw < Thresh1 ~ 1,
      Thresh1_basis %in% "D" & perc90_value_dw >= Thresh1 & is.na(Thresh2_basis) ~ 2,
      Thresh1_basis %in% "D" & perc90_value_dw >= Thresh1 & Thresh2_basis %in% "W" & perc90_value_ww < Thresh2 ~ 2,
      Thresh1_basis %in% "D" & perc90_value_dw >= Thresh1 & Thresh2_basis %in% "W" & perc90_value_ww >= Thresh2 ~ 3,
      Thresh1_basis %in% "D" & perc90_value_dw >= Thresh1 & Thresh2_basis %in% "L" & perc90_value_fw < Thresh2 ~ 2,
      Thresh1_basis %in% "D" & perc90_value_dw >= Thresh1 & Thresh2_basis %in% "L" & perc90_value_fw >= Thresh2 ~ 3
      )
  ) 


xtabs(~PARAM + Status, dat_status_2)

param <- "HG"
cat("\nStatus for", param, ":\n")
xtabs(~Country + Status, dat_status_2 %>% filter(PARAM %in% param))
  
if (FALSE){
  # for sensibility check
  dat_status_2 %>% distinct(PARAM, BASIS, Thresh1, Thresh1_basis, Thresh1_as) 
  dat_status_2 %>% distinct(PARAM, BASIS, Thresh2, Thresh2_basis, Thresh2_as) 
}


```

### dat_status_3
```{r}

dat_status_3 <- dat_series %>%
  left_join(dat_status_2, by = c("Country", "id", "WoRMS_scientificName", "PARAM"))

xtabs(~addNA(Status), dat_status_3)


```


### For debugging  
```{r}

if (FALSE){
  
  test <- dat_status_2 %>% 
    mutate(
      Status2 = case_when(
        Thresh1_basis %in% "W" & perc90_value_ww < Thresh1 ~ 1,
        Thresh1_basis %in% "W" & perc90_value_ww >= Thresh1 & is.na(Thresh2_basis) ~ 2,
        Thresh1_basis %in% "W" & perc90_value_ww >= Thresh1 & Thresh2_basis %in% "W" & perc90_value_ww < Thresh2 ~ 3,
        Thresh1_basis %in% "W" & perc90_value_ww >= Thresh1 & Thresh2_basis %in% "W" & perc90_value_ww >= Thresh2 ~ 4,
        Thresh1_basis %in% "W" & perc90_value_ww >= Thresh1 & Thresh2_basis %in% "L" & perc90_value_fw < Thresh2 ~ 5,
        Thresh1_basis %in% "W" & perc90_value_ww >= Thresh1 & Thresh2_basis %in% "L" & perc90_value_fw >= Thresh2 ~ 6,
        Thresh1_basis %in% "D" & perc90_value_dw < Thresh1 ~ 7,
        Thresh1_basis %in% "D" & perc90_value_dw >= Thresh1 & is.na(Thresh2_basis) ~ 2,
        Thresh1_basis %in% "D" & perc90_value_dw >= Thresh1 & Thresh2_basis %in% "W" & perc90_value_ww < Thresh2 ~ 8,
        Thresh1_basis %in% "D" & perc90_value_dw >= Thresh1 & Thresh2_basis %in% "W" & perc90_value_ww >= Thresh2 ~ 9,
        Thresh1_basis %in% "D" & perc90_value_dw >= Thresh1 & Thresh2_basis %in% "L" & perc90_value_fw < Thresh2 ~ 10,
        Thresh1_basis %in% "D" & perc90_value_dw >= Thresh1 & Thresh2_basis %in% "L" & perc90_value_fw >= Thresh2 ~ 11,
        TRUE ~ 12
      )
  ) 
  
  
  # dry weight
  param <- "CB118"
  test %>%
    filter(PARAM == param) %>%
    select(PARAM, perc90_value_dw, perc90_value_ww, perc90_value_fw, 
           Thresh1, Thresh1_basis, Thresh2, Thresh2_basis, Status, Status2) %>%
    View(param)

  param <- "CB118"
  test %>%
    filter(PARAM == param) %>%
    ggplot(aes(perc90_value_dw, perc90_value_fw, color = factor(Status))) +
    geom_point() +
    scale_x_log10() +
    scale_y_log10() +
    geom_vline(aes(xintercept = 0.60)) +
    geom_hline(aes(yintercept = 25))
    
  
    }
```



### test plot 1, threshold level 1       
```{r}

source("06_MAR001_results_2021_functions.R")

# debugonce(plot_level_vs_species)
plot_level_vs_species("CD", data = dat_status_3)
plot_level_vs_species("CU", data = dat_status_3)
plot_level_vs_species("PB", data = dat_status_3)
plot_level_vs_species("HG", data = dat_status_3)
plot_level_vs_species("HCB", data = dat_status_3)    
plot_level_vs_species("HCHG", data = dat_status_3)    
plot_level_vs_species("DDEPP", data = dat_status_3)
plot_level_vs_species("CB118", data = dat_status_3)     
plot_level_vs_species("BAP", data = dat_status_3)



```

### test plot 2: threshold level 2       
```{r}

# debugonce(plot_level_vs_species)
plot_level_vs_species("HG", threshold = 2, data = dat_status_3)
plot_level_vs_species("HCB", threshold = 2, data = dat_status_3)  
plot_level_vs_species("HCHG", threshold = 2, data = dat_status_3)  
plot_level_vs_species("CB118", threshold = 2, data = dat_status_3)  

# param <- "CB118"
# dat_status_3 %>% filter(PARAM == param) %>% View(param)


```

### test plot 2   
```{r}

param <- "HCHG"

dat_status_3 %>%
  filter(PARAM == param) %>%    #  & Country == "Norway"
  ggplot(aes(Longitude, Latitude, color = factor(Status))) +
  geom_point() +
  scale_color_manual(values = c(`1` = "blue", `2` = "green", `3` = "red"))

```

## 8. Combine status and trend    

### Combine by joining
- dat_series + result_trends + dat_status_2

```{r, results='hold'}

### dat_status_3

names(dat_series) %>% paste(collapse = ", ")
names(dat_series_trend) %>% paste(collapse = ", ")
names(dat_status_2) %>% paste(collapse = ", ")

#
# Check 'dat_series_trend'  
# 
check <- dat_series_trend %>%
  count(Country, id, WoRMS_scientificName, PARAM) %>%
  filter(n > 1) %>%
  nrow()
if (check > 0)
  stop("Check whether 'dat_series_trend' has unique rows")

dat_status_trend <- dat_series %>%
  left_join(
    dat_series_trend %>% 
      select(Country, id, WoRMS_scientificName, PARAM, Slope, trend), , 
    by = c("Country", "id", "WoRMS_scientificName", "PARAM")) %>%
  mutate(
    trend = ifelse(is.na(trend), "Trend unknown", trend)) %>%
  left_join(
    dat_status_2 %>%
      filter(status_analysis %in% "Yes") %>%
      select(Country, id, WoRMS_scientificName, PARAM,
             Status, status_analysis,
             perc90_value_ww, perc90_value_dw, perc90_value_fw, Group, 
             Thresh1, Thresh1_basis, Thresh2, Thresh2_basis, Th1, Th2), 
    by = c("Country", "id", "WoRMS_scientificName", "PARAM")) %>%
  rename(Trend = trend) %>%
  filter(!is.na(Status) & status_analysis %in% "Yes") %>%
  mutate(
    Region = case_when(
      MSFD_region %in% "BAL" ~ "Baltic",
      MSFD_region %in% c("MWE","MAD","MIC") ~ "Mediterranean",
      MSFD_region %in% "BLK" ~ "Black Sea",
      TRUE ~ "North-East Atlantic Ocean"
    )
  )
  
xtabs(~Status + Trend, dat_status_trend)


```


### Save  
```{r}

# Save backup with time stamp
filename <- "Data/07_dat_status_trend.rds"
filename_backup <- sub(".rds", paste0("_", lubridate::now(), ".rds"), filename, fixed = TRUE)
filename_backup <- gsub(":", "-", filename_backup, fixed = TRUE)
filename_backup <- gsub(" ", "_", filename_backup, fixed = TRUE)
filename_backup
file.copy(filename, filename_backup)

# UNCOMMENT to save
saveRDS(dat_status_trend, filename)

# NOTE: version before the change on 2022-06-17 (commit 1426b98, see text at the top of this script)
#   was saved as '07_dat_status_trend_BEFORE_UPDATE_2022-06-17.rds'  
#   and used in a special version of 

# Read back?
# dat_status_trend <- readRDS("Data/07_dat_status_trend.rds")

# Old version (using medians)
# saveRDS(dat_status_old, "Data/05_dat_status_trend_OLD.rds")

xtabs(~PARAM + Status, dat_status_trend)

```

## 9. Extras  

### Check for HG limits  
```{r, results='hold'}

# Individual data  

cat("----------------------------------------------------------------------------  \n")
cat("dat_ind (data for individual samples), ww basis compared with 20, dw basis compared with 129   \n")
cat("----------------------------------------------------------------------------  \n")

dat_ind %>%
  filter(MYEAR >= 2010,
         PARAM == "HG", 
         Country %in% c("Germany", "Denmark", "Sweden", "Poland"),
         BASIS == "W") %>%
  xtabs(~(Value > 20) + Country + BASIS, .)

dat_ind %>%
  filter(MYEAR >= 2010,
         PARAM == "HG", 
         Country %in% c("Germany", "Denmark", "Sweden", "Poland"),
         BASIS == "D") %>%
  xtabs(~(Value > 129) + Country + BASIS, .)  

cat("----------------------------------------------------------------------------  \n")
cat("dat_medians (medians per year), 'Value' is dw basis in case of mercury    \n")
cat("----------------------------------------------------------------------------  \n")

dat_medians %>%
  filter(Year >= 2010,
         PARAM == "HG", 
         Country %in% c("Germany", "Denmark", "Sweden", "Poland")) %>%
  xtabs(~(Value > 129) + Country, .)  

cat("----------------------------------------------------------------------------  \n")
cat("dat_status_3 (status)   \n")
cat("----------------------------------------------------------------------------  \n")

dat_status_3 %>%
  filter(PARAM == "HG", 
         Country %in% c("Germany", "Denmark", "Sweden", "Poland")) %>%
  xtabs(~Status + Country + status_analysis, .)  

cat("----------------------------------------------------------------------------  \n")
cat("dat_status_3 (status) by region (all countries)  \n")
cat("----------------------------------------------------------------------------  \n")

dat_status_3 %>%
  filter(PARAM == "HG") %>%
  xtabs(~Status + Region + status_analysis, .)  

cat("----------------------------------------------------------------------------  \n")
cat("dat_status_trend  \n")
cat("----------------------------------------------------------------------------  \n")

dat_status_trend %>%
  filter(PARAM == "HG", 
         Country %in% c("Germany", "Denmark", "Sweden", "Poland")) %>% # View()
  xtabs(~Status + Country, .)  


```

### Check HG status (table and map)   
```{r}

if (F){
  
  #
  # View HG status data as table
  #
  dat_status_3 %>% filter(PARAM == "HG") %>% View("dat_status_3")
  
  #
  # View HG status data on a map (status_analysis = Yes only)
  #
  # Saved as
  #   '07_map_Hg_status_this_script.JPG'  
  # Compared with OSPAR OHAT map (2022 version)
  #   https://dome.ices.dk/ohat/?assessmentperiod=2022
  #   '07_map_Hg_status_OSPAR_OHAT.JPG'
  
  #
  df1 <- dat_status_3 %>%
    filter(PARAM == "HG", status_analysis %in% "Yes", Status == 1)
  df2 <- dat_status_3 %>%
    filter(PARAM == "HG", status_analysis %in% "Yes", Status == 2)
  df3 <- dat_status_3 %>%
    filter(PARAM == "HG", status_analysis %in% "Yes", Status == 3)
  
  icons_green <- awesomeIcons(
    icon = 'ios-close', iconColor = 'black', library = 'ion', markerColor = "green"
  )
  icons_red <- awesomeIcons(
    icon = 'ios-close', iconColor = 'black', library = 'ion', markerColor = "red"
  )
  
  leaflet() %>%
    addTiles() %>%
    addAwesomeMarkers(lng = df2$Longitude, lat = df2$Latitude, popup = df2$id, icon = icons_red) %>%
    addAwesomeMarkers(lng = df1$Longitude, lat = df1$Latitude, popup = df1$id, icon = icons_green)
  
}

```


### Check of regions, using mapview         

```{r}

dat_assessed_stations <- sf::st_as_sf(
  dat_status_trend %>% 
    group_by(Country, Region, MSFD_region, id, Longitude, Latitude) %>%
    summarise(
      species = paste(unique(WoRMS_scientificName), collapse = ", "),
      n_species = length(unique(WoRMS_scientificName)),
      first_year = min(first_year),
      last_year = max(last_year),
      n_years = max(n_years)
    ) %>%
    mutate(Lon = Longitude, Lat = Latitude),  # copy these, so they show up in the pop-up
  coords = c("Longitude", "Latitude"),
  crs = "+proj=longlat +ellps=WGS84",
  agr = "constant")

# in console:
# mapview::mapview(dat_assessed_stations, zcol = "MSFD_region")
# mapview::mapview(dat_assessed_stations, zcol = "Region")

```



### mapview check    
- Compare with https://odims.ospar.org/en/submissions/ospar_outer_boundary_2016_01/ 
```{r}

sf_CD <- sf::st_as_sf(
  dat_status_trend %>% 
    filter(PARAM == "CD") %>%
    mutate(Lon = Longitude, Lat = Latitude),  # copy these, so they show up in the pop-up
  coords = c("Longitude", "Latitude"),
  crs = "+proj=longlat +ellps=WGS84",
  agr = "constant")

# in console:
# mapview::mapview(sf_CD, zcol = "id")

# head(sf_CD)

```


### test plot  
```{r, fig.width=9, fig.height=7}


library(maps)
my_map <- map_data("world")

param <- "HCHG"
param <- "HG"
# param <- "CB118"

df_plot <- dat_status_trend %>%
  mutate(Status = factor(Status)) %>%
  filter(PARAM == param)    #  & Country == "Norway"

ggplot(df_plot, aes(Longitude, Latitude, fill = Status, shape = Trend)) +
  annotation_map(my_map, fill = "grey70") +
  geom_point() +
  scale_fill_manual(values = c(`1` = "blue", `2` = "green", `3` = "red")) +
  scale_shape_manual(values = c(Decrease = 25, Increase = 24, 
                                `No trend` = 21, `Trend unknown` = 22)) +
  coord_map("ortho", orientation = c(52, 10, 0))


```


## Appendix  


### From 4a -  Get regression results, using list columns  
The tidy way - stopped working, probably because {maps} was loaded, stopping map() from working   
Made lots of mess trying to fix this (before realizing the reason), should be cleaned  
```{r}

if (FALSE){
  
  str(m, 2)

  dat_medians_ols$data[2]
  
  result_ols_model_all1 <- dat_medians_ols %>%
    mutate(
      ols_model = map(data, lm_fixed_s)    # perform lm
    )
  
  not_ok <- map_lgl(result_ols_model_all1$ols_model, ~is.null(.x[[1]]))
  cat(sum(!not_ok), "of", length(not_ok), "results are ok \n")
  
  # result_ols_model_all$ols_model[1] %>% broom::tidy()
  # result_ols_model_all$ols_model[1][[1]] %>% broom::tidy()
  # result_ols_model_all$ols_model[1][[1]] %>% broom::tidy()
  # result_ols_model_all$ols_model[1][[1]]$result %>% broom::tidy()
  
 
  result_ols_model_all <- dat_medians_ols %>%
    mutate(
      ols_model = map_peacefully(data, ~lm(Log_value ~ Year, data = .x)),    # perform lm
      ols_coef = map_peacefully(ols_model, ~broom::tidy(.$x[[1]]))    
    )
  
  
  
  not_ok <- map_lgl(result_ols_model_all$ols_model, ~is.null(.x[[1]]))
  cat(sum(!not_ok), "of", length(not_ok), "results are ok \n")
  
  
  #    ols_coef = map_peacefully(ols_model, tidy)                          
  
  result_ols <- result_ols_model %>%
    select(-ols_model) %>%
    unnest(ols_coef) %>% 
    filter(term == "Year") %>%
    mutate(
      trend = case_when(
        estimate > 0 & p.value < 0.05 ~ "Increase",
        estimate < 0 & p.value < 0.05 ~ "Decrease",
        TRUE ~ "No trend")
    )
  

xtabs(~trend, result_ols)
xtabs(~Country + trend, result_ols)
xtabs(~PARAM + trend, result_ols)


# dat_medians_ols$mod_ols[1]
# dat_medians_ols[1,]  

}

```

