---
title: "11_export_data"
format: html
---


Export data to be submitted to the consortium   

* 
* Note: excel for the EQS ratio map for CB118 (`06_Relative_class_CB118_by_MSFDregion_plotdata.xlsx`) was made in script 06   


## Libraries  
```{r, echo = FALSE, message=FALSE, warning=FALSE, results='hide'}

library(dplyr)  
library(ggplot2)

```


## Data  
```{r data, echo=FALSE, results='hold'}

#
# Simple categorical trend + status
#

dir("Data", "07_dat_status_trend_2024_")
fn <- "Data/07_dat_status_trend_2024_2024-10-23_14-11-04.091695.rds"
  
dat_status_trend <- readRDS(fn)
# nrow(dat_status_trend)

# Delete one PCB118 outlier
dat_status_trend <- dat_status_trend %>%
  filter(!(id %in% "56.02833333_20.83333333" & PARAM %in% "CB118")) %>%
  mutate(
    Status = case_when(
      Status == 1 ~ "Low",
      Status == 2 ~ "Moderate",
      Status == 3 ~ "High")
  )
# nrow(dat_status_trend)

cat("File modified at: \n")
file.info(fn)$mtime

#
# Meta-analysis of trends  
#
meta_trends_region <- readRDS("Data/06_meta_trends_region.rds")


```

## Export data for maps  
```{r}

# dir.create("Submitted")

# write.csv(dat_status_trend, "Submitted/2024/MAR001_map_data_points.csv")  
write.csv(dat_status_trend, "Submitted_to_EEA/2024/MAR001_map_data_points.csv")  

writeLines(
  c(
    "This file contains data needed for making maps for each contaminant. ",
    "",
    "The most important information for this is (apart from the coordinates) the columns Trends and Status. ",
    "Status has the values 1, 2 and 3 (low, moderate and high concentrations) and can be plotted using the colors",
    "green, yellow and red"
  ), 
  "Submitted_to_EEA/2024/MAR001_map_data_points_info.csv")  


```


## Data for pie diagrams  

### Status by region

```{r, fig.width=10, fig.height=4}

status_by_region <- dat_status_trend %>%
  mutate(PARAM = factor(
    PARAM, 
    levels = c("CD","CU","PB","HG","HCB","HCHG","DDEPP","BAP", "CB118")),
    Status = factor(Status, levels = c("Low", "Moderate", "High"))
  ) %>%
  count(Region, PARAM, Status, .drop = FALSE) %>%
  group_by(PARAM, Region) %>%
  mutate(
    Total = sum(n)
  ) %>%
  ungroup() %>%
  mutate(
    `Percent (for parameter/region)` = round(100*n/Total,1)
  )

```


### Trends by region   

* Not included in finbal excel  

```{r, fig.width=10, fig.height=4}

trend_by_region <- dat_status_trend %>%
  mutate(PARAM = factor(
    PARAM, 
    levels = c("CD","CU","PB","HG","HCB","HCHG","DDEPP","BAP", "CB118"))
  ) %>%
  count(PARAM, Region, Trend) %>%
  group_by(PARAM, Region) %>%
  mutate(
    Total = sum(n)
  ) %>%
  ungroup() %>%
  mutate(
    `Percent (for parameter/region)` = round(100*n/Total,1)
  )
  

```


### Meta-analysis of trends  

```{r}

# names(meta_trends_region) %>% paste(collapse = ", ")

trend_by_region_meta <- meta_trends_region %>%
  select(PARAM, Region, Change_perc_10yr, Change_perc_10yr_lo, Change_perc_10yr_up, 
         overall_trend = `Overall trend`) %>%
  filter(PARAM %in% unique(status_by_region$PARAM))


```


### Save in excel  

#### Parameter names to use in output  

* See https://www.eea.europa.eu/en/analysis/indicators/hazardous-substances-in-marine-organisms

```{r}

unique(dat_status_trend$PARAM) %>% dput()
unique(dat_status_trend$Status) %>% dput()
unique(dat_status_trend$Trend) %>% dput()
unique(trend_by_region_meta$`Overall trend`) %>% dput()

lookup_param_mapname <- data.frame(
  PARAM = c("BAP", "CB118", "CD", "CU", "DDEPP", "HCB", "HCHG", "HG", "PB"),
  map_name = c("B(a)P", "PCB118", "Cd", "Cu", "p,p'-DDE", "HCB", "HCHG", "Hg", "Pb")
)

lookup_arrow_direction <- data.frame(
  overall_trend = c("No sign. pattern", "Downward", "Upward"),
  arrow_direction = c("Towards the right", "Downward", "Upward"),
  legend = c("No change", "Improvement (decreasing concentrations)", 
             "Detoriation (increasing concentrations)")
)

points_on_map <- dat_status_trend %>% 
  left_join(lookup_param_mapname, by = join_by(PARAM), relationship = "many-to-one") %>% 
  select(map_name, Longitude, Latitude, Concentrations = Status, `Concentration trend` = Trend)

pie_plots <- status_by_region %>% 
  left_join(lookup_param_mapname, by = join_by(PARAM), relationship = "many-to-one") %>% 
  select(map_name, Region, Status, Percentage = `Percent (for parameter/region)`)

pie_plots_arrows <- trend_by_region_meta %>% 
  left_join(lookup_param_mapname, by = join_by(PARAM), relationship = "many-to-one") %>%
  left_join(lookup_arrow_direction, by = join_by(overall_trend), relationship = "many-to-one") %>% 
  select(map_name, Region, arrow_direction, legend)

colours <- data.frame(
  c("High", "Moderate", "Low"),
  c("Red", "Yellow", "Green")
)


```
#### Make and save excel data  

```{r}

data_list <-   list(
  "Points on map" = points_on_map,
  "Pie plots" = pie_plots, 
  "Colours (map and pies)" = colours, 
  "Pie plots arrows" = pie_plots_arrows,
  "Info" = data.frame(
    Sheet = c(
      "Points on map",
      "Pie plots", 
      "Pie plots arrows"),
    Info = c(
      "Points on Europe map, with shape (column 'Concentration trend') and colour (column Concentration)",
      "Pie plots for each region. Percentages sum to 100% for each combintion of parameter and region",
      "Regional concentration trend, shown as arrows next to pies (not enough data for Mediterranean pies)"
    )
  )
)

str(data_list, 1)

writexl::write_xlsx(data_list, "Submitted_to_EEA/2024/MAR001_DataPackageFig1_before_edit.xlsx")


```

## Test graphs  

### status_by_region  

```{r, fig.width=9, fig.height=3}

ggplot(status_by_region, aes(x = PARAM, y = `Percent (for parameter/region)`, fill = Status)) +
  geom_col() +
  scale_fill_manual(values = c("green3", "yellow", "red3")) +
  facet_wrap(vars(Region))


```

### status_by_region, pie plot  

```{r}

ggplot(status_by_region, aes(x = "", y = `Percent (for parameter/region)`, fill = Status)) +
  geom_col() +
  scale_fill_manual(values = c("green3", "yellow", "red3")) +
  coord_polar(theta = "y") +
  facet_grid(vars(Region), vars(PARAM))

```

### trend_by_region  

```{r, fig.width=9, fig.height=3}

# ggplot(trend_by_region, aes(x = PARAM, y = `Percent (for parameter/region)`, fill = factor(Trend))) +
#   geom_col() +
#   scale_fill_manual(values = c("green3", "red3", "grey60", "grey30")) +
#   facet_wrap(vars(Region))

```


```{r, fig.width=9, fig.height=7}

ggplot(trend_by_region_meta, aes(x = Region, y = Change_perc_10yr)) +
  geom_pointrange(aes(ymin = Change_perc_10yr_lo, ymax = Change_perc_10yr_up, color = `Overall trend`)) +
  geom_hline(yintercept = 0) +
  scale_colour_manual(values = c("Downward" = "green3", "No sign. pattern" = "grey30")) +
  coord_flip(ylim = range(trend_by_region_meta$Change_perc_10yr)) +
  facet_wrap(vars(PARAM))  

```
