---
title: "32_PFAS_2025"
author: "DHJ"
date: "2025-09-02"
output: 
  html_document:
    keep_md: true  
    toc: true  
    toc_float: true
---

## 0. Settings
```{r}

save_plots <- FALSE

```


## 1. Packages    
```{r, message=FALSE, warning=FALSE, results='hide'}

library(dplyr)
library(purrr)
library(lubridate)
library(readr)
library(stringr)    
library(ggplot2)
library(ggeasy)        # easy formatting of ggplot plots
library(forcats)       # easy ordering of levels in a factor
library(RColorBrewer)  
library(glue)

library(agricolae)


```

## 2. Read ICES data

*  No PFAS in EMODnet data   


### ICES data

```{r}

# As EMODnet, only that L20 and botDepth 
coltypes <- cols(
  .default = col_character(),
  MYEAR = col_integer(),
  SD_StationCode = col_double(),
  Samplingtime = col_date(format = ""),
  Latitude = col_double(),
  Longitude = col_double(),
  DEPHU = col_double(),
  DEPHL = col_double(),
  SampleIdentifier = col_character(),
  SubsampleIdentifier = col_character(),
  WoRMS_AphiaID = col_character(),
  WoRMS_accepted_AphiaID = col_character(),
  SEXCO = col_character(),
  Value = col_double(),
  # L20 = col_double(),
  UNCRT = col_double(),
  LMQNT = col_double(),
  DETLI = col_double(),
  DRYWT = col_double(),
  FATWT = col_double(),
  EXLIP = col_double(),
  LIPIDWT = col_double()
  # BotDepth = col_double()
)

df2_orig <- readr::read_tsv("Input_data/From_Hans_2025/ICES_DOME_biota_20250625.txt", col_types = coltypes) %>%
  mutate(Year = year(Samplingtime), 
         Month = month(Samplingtime), 
         .before = everything())

nrow(df2_orig)
# 1635146


```


## 3. Define lookup terms  

### Define species type and common name  

* ICES data  

```{r}

# For writing code:
# df2_orig %>%
#   filter(PARAM == "PFOS") %>% pull(WoRMS_scientificName) %>% unique() %>% dput

df_taxon <- tribble(
  ~WoRMS_scientificName, ~Species_type, ~Common_name,
  "Clupea harengus", "Fish, low-trophic", "Herring", 
  "Uria aalge", "Seabird", "Common guillemot", 
  "Platichthys flesus", "Fish, benthic", "European flounder",
  "Pleuronectes platessa", "Fish, benthic", "European plaice", 
  "Crassostrea gigas", "Mussel/oyster", "Pacific oyster",
  "Mytilus edulis", "Mussel/oyster", "Blue mussel",
  "Gadus morhua", "Fish, predatory", "Cod",
  "Perca fluviatilis", "Fish, predatory", "European perch",
  "Haematopus ostralegus", "Seabird", "Oystercatcher",
  "Sterna hirundo", "Seabird", "Common tern",
  "Macoma balthica", "Mussel/oyster", "Baltic macoma",
  "Zoarces viviparus", "Fish, low-trophic", "Eelpout", 
  "Neogobius melanostomus", "Fish, benthic", "Round goby", 
  "Somateria mollissima", "Seabird", "Eider duck",
  "Ursus maritimus", "Mammal", "Polar bear", 
  "Phoca hispida", "Mammal", "Ringed seal", 
  "Globicephala melas", "Mammal", "Pilot whale", 
  "Ovis aries", "Terrestrial mammal", "Sheep",
  "Abramis brama", "Fish, low-trophic", "Bream", 
  "Fulmarus glacialis", "Seabird", "Northern fulmar",
  "Cepphus grylle", "Seabird", "Black guillemot"
)

```

### Define region names   

```{r}


if (FALSE){
  ggplot(df2_orig %>% distinct(Longitude, Latitude, MSFD_region), 
         aes(Longitude, Latitude, color = MSFD_region)) +
    geom_point()
}


df_region <- tribble(
  ~MSFD_region, ~MSFD_regionname,
  "ABI", "Biscay, Iberian coast",
  "ACS", "Celtic Seas",
  "ANS", "Greater North Sea",
  "ATL", "Atlantic Ocean",
  "BAL", "Baltic Sea",
  "BAR", "Barents Sea",
  "NOR", "Norwegian Sea"
)


```


### Add species type and common names + region  

```{r}

df2_orig2 <- df2_orig %>%
  left_join(df_taxon, by = "WoRMS_scientificName") %>%
  left_join(df_region, by = "MSFD_region")


```

## 4. Tables  

### Number of measurements   

* ICES only, as there are no EMODnet data   

```{r}

df2_orig2 %>%
  filter(PARGROUP == "O-FL") %>%
  xtabs(~ MYEAR + PARAM, .)

df2_orig2 %>%
  filter(PARGROUP == "O-FL") %>%
  group_by(Country, STATN, PARAM) %>%
  summarize(
    n_year = n(),
    min_year = min(MYEAR),
    max_year = max(MYEAR) 
  ) %>%
  filter(
    max_year >= 2011   
  ) %>%
  xtabs(~ Country + PARAM, .)


```



### Number of series  

```{r}

df_pfas_series <- df2_orig2 %>%
  filter(PARGROUP == "O-FL") %>%
  distinct(Country, MSFD_regionname, STATN, PARAM, WoRMS_scientificName, Common_name, MATRX, MYEAR) %>%
  group_by(Country, MSFD_regionname, STATN, PARAM, WoRMS_scientificName, Common_name, MATRX) %>%
  summarize(
    n_year = n(),
    min_year = min(MYEAR),
    max_year = max(MYEAR) 
  ) %>%
  filter(
    max_year >= 2011   
  )

xtabs(~ Common_name + PARAM, df_pfas_series)

```


### Species vs regions  

```{r}

xtabs(~Common_name + MSFD_regionname, df_pfas_series)

```

## 5. PFOS data   

### Plot series     

```{r, fig.width = 6, fig.height=12}

df_pfas_series %>%
  ungroup() %>%
  filter(PARAM == "PFOS",
         max_year > min_year) %>%
  arrange(desc(max_year), desc(min_year)) %>%
  mutate(
    Series = fct_inorder(factor(paste(Country, STATN)))
    ) %>%
  ggplot(aes(x = Series, ymin = min_year, ymax = max_year)) +
  geom_linerange() +
  coord_flip()

```

### Species colors and order  

```{r}

# RColorBrewer::display.brewer.all(5)

cols_purple <- RColorBrewer::brewer.pal(5, name = "Purples")[4]
cols_blue <- RColorBrewer::brewer.pal(5, name = "Blues")[3:5]
cols_red <- RColorBrewer::brewer.pal(5, name = "OrRd")[c(3,5)]
cols_green <- RColorBrewer::brewer.pal(5, name = "Greens")[c(3,5)]

species_type_colors <- c("Mussel/oyster" = cols_purple, 
                        "Fish, low-trophic" = cols_blue[1],
                        "Fish, benthic" = cols_blue[2],
                        "Fish, predatory" = cols_blue[3], 
                        "Seabird" = cols_red[1],
                        "Mammal" = cols_red[2])

species_type_order <- names(species_type_colors)

```

### Region order  

```{r}

region_colors <- c("Barents Sea" = cols_purple, 
                   "Atlantic Ocean" = cols_blue[1],
                   "Norwegian Sea" = cols_blue[2],
                   "Celtic Seas" = cols_blue[3], 
                   "Greater North Sea" = cols_red[1],
                   "Baltic Sea" = cols_red[2],
                   "Biscay, Iberian coast" = cols_green[2])

region_order = names(region_colors)

```

### Basis and units     

```{r}
df2_orig2 %>%
  filter(PARAM == "PFOS") %>%
  xtabs(~MUNIT + BASIS, .)

df2_orig2 %>%
  filter(PARAM == "PFOS") %>%
  xtabs(~WoRMS_scientificName + BASIS, .)


```



### PFOS data  

```{r}

df_pfos <- df2_orig2 %>%
  filter(PARAM == "PFOS" & BASIS == "W" & Species_type != "Terrestrial mammal") %>%
  mutate(
    Value = case_when(
      MUNIT == "ng/kg" ~ Value/1000,
      TRUE ~ Value),
    Conc_EQS_ratio = Value/9.1,
    Species_type = factor(Species_type, levels = species_type_order),
    MSFD_regionname = factor(MSFD_regionname, levels = region_order)
    ) %>%
  mutate(Common_name = fct_reorder(factor(Common_name), Value))

```

## 6. PFOS plots   

### Ratio concentration/EQS    

```{r}

# display.brewer.all(n = 5)


gg <- ggplot(df_pfos, aes(x = Common_name, y = Conc_EQS_ratio, color = Species_type)) +
  geom_hline(yintercept = 1) +
  geom_boxplot() +
  scale_color_manual("Species group", values = species_type_colors) +
  scale_y_log10(labels = scales::label_number()) +
  theme_bw() +
  easy_rotate_labels("x", angle = -45) +
  easy_remove_axes(which = "x", what = "title") +
  labs(y = "Ratio concentration/EQS")

if (save_plots)
  ggsave("Figures/2025/32_PFAS_EQSratio1.png", gg, width = 8, height = 5, dpi = 200)

gg

```
#### Get raw data for plot  

```{r}

df_pfos_for_plot <- df_pfos %>%
  select(Common_name, Conc_EQS_ratio, Species_type)

# Saved further down (next chunk)

# Test
gg <- ggplot(df_pfos_for_plot, aes(x = Common_name, y = Conc_EQS_ratio, color = Species_type)) +
  geom_hline(yintercept = 1) +
  geom_boxplot() +
  scale_color_manual("Species group", values = species_type_colors) +
  scale_y_log10(labels = scales::label_number()) +
  theme_bw() +
  easy_rotate_labels("x", angle = -45) +
  easy_remove_axes(which = "x", what = "title") +
  labs(y = "Ratio concentration/EQS")

# gg

```

#### Get plot data  

```{r}
ggplot_build(gg) %>% str(1)
ggplot_build(gg)$data %>% str(1)
ggplot_build(gg)$data[[2]]

a <- layer_data(gg, i = 2)
str(a, 1)

a2 <- a %>%
  mutate(colour = fct_inorder(colour)) %>%
  mutate(Common_name = levels(df_pfos$Common_name)[x],
         Common_name = fct_inorder(Common_name))

# Boxes
a2_boxes <- a2 %>% 
  select(x, Common_name, ymin, lower, middle, upper, ymax, colour)

# - numbers on x axis
ggplot(a2_boxes, aes(x, middle, color = colour)) +
  geom_point(pch = 15) +
  geom_point(aes(y = ymin), pch = 1) +
  geom_point(aes(y = lower), pch = 6) +
  geom_point(aes(y = upper), pch = 2) +
  geom_point(aes(y = ymax), pch = 1) +
  scale_color_manual(values = levels(a2$colour))

# - names on x axis
ggplot(a2_boxes, aes(Common_name, middle, color = colour)) +
  geom_point(pch = 15) +
  geom_point(aes(y = ymin), pch = 1) +
  geom_point(aes(y = lower), pch = 6) +
  geom_point(aes(y = upper), pch = 2) +
  geom_point(aes(y = ymax), pch = 1) +
  scale_color_manual(values = levels(a2$colour)) +
  easy_rotate_labels("x", angle = -45) 

# Outliers  
a2_outliers <- a2 %>% 
  select(Common_name, x, colour, outliers) %>%
  tidyr::unnest(cols = c(outliers))
ggplot(a2_outliers, aes(x, outliers, color = colour)) +
  geom_point(pch = 15) +
  scale_color_manual(values = levels(a2$colour))

```

#### Save both types  

```{r}

# Save
writexl::write_xlsx(
  list(points = a2_outliers,
       boxplot = a2_boxes %>% 
         rename(
           `Lower end of whiskers` = ymin,
           `Upper end of whiskers` = ymax),
       raw_data = df_pfos_for_plot,
       colours = data.frame(
         species_group = names(species_type_colors),
         colors = species_type_colors)),
  "Figures/2025/PFOS_signal_data.xlsx" 
)

```


#### With regions  

```{r}

# xtabs(~MSFD_regionname, df_pfos)

# dput(unique(df_pfos$MSFD_regionname))

gg2 <- gg + 
  facet_wrap(vars(MSFD_regionname))  

if (save_plots)
  ggsave("Figures/2025/32_PFAS_EQSratio2.png", gg2, width = 9, height = 8, dpi = 200)

gg2

```

### PFOS in cod  

#### Boxplot, all years  

```{r}

species <- "Cod"

df_pfos_cod <- df_pfos %>%
  filter(Common_name == species)

gg <- ggplot(df_pfos_cod, aes(x = MSFD_regionname, y = Conc_EQS_ratio, color = MSFD_regionname)) +
  geom_hline(yintercept = 1) +
  geom_boxplot() +
  scale_color_manual(values = region_colors) +
  scale_y_log10(labels = scales::label_number()) +
  theme_bw() +
  easy_rotate_labels("x", angle = -45) +
  easy_remove_axes(which = "x", what = "title") +
  labs(y = "Ratio concentration/EQS",
       title = paste("PFOS in", tolower(species)))

fn <- glue("Figures/2025/32_PFAS_EQSratio_{tolower(species)}.png")
if (save_plots)
  ggsave(fn, gg, width = 8, height = 5, dpi = 200)

gg

```

#### Boxplot, 2015-2016 and 2021 only    

```{r}

df_pfos_cod_sel <- df_pfos_cod %>%
  mutate(
    Years = case_when(
      Year %in% 2015:2016 ~ "2015-2016",
      Year %in% 2021 ~ "2021")
    ) %>%
  filter(!is.na(Years))

gg <- ggplot(df_pfos_cod_sel, aes(x = MSFD_regionname, y = Conc_EQS_ratio, color = MSFD_regionname)) +
  geom_hline(yintercept = 1) +
  geom_boxplot() +
  scale_color_manual(values = region_colors) +
  scale_y_log10(labels = scales::label_number()) +
  theme_bw() +
  easy_rotate_labels("x", angle = -45) +
  easy_remove_axes(which = "x", what = "title") +
  labs(y = "Ratio concentration/EQS",
       title = paste("PFOS in", tolower(species))) +
  facet_grid(cols = vars(Years))

#fn <- glue("Figures/2025/32_PFAS_EQSratio_{tolower(species)}.png")
# ggsave(fn, gg, width = 8, height = 5, dpi = 200)

gg

```

#### Data by year  
```{r}

df_pfos_cod %>%
  count(Year, MSFD_regionname) %>%
  ggplot(aes(Year, MSFD_regionname, fill = n)) +
  geom_tile()

```


#### Test effect of region and year (ANOVA)    

```{r}

nrow(df_pfos_cod)

df_pfos_cod <- df_pfos_cod %>%
  mutate(log_Conc_EQS_ratio = log10(Conc_EQS_ratio))

m0 <- lm(log_Conc_EQS_ratio ~ 1, data = df_pfos_cod)
m1a <- lm(log_Conc_EQS_ratio ~ MSFD_regionname, data = df_pfos_cod)
m1b <- lm(log_Conc_EQS_ratio ~ Year, data = df_pfos_cod)
m2 <- lm(log_Conc_EQS_ratio ~ MSFD_regionname + Year, data = df_pfos_cod)
cat("--------------------------------------------\n")
cat("Null model -> Region -> Region + Year \n")
cat("--------------------------------------------\n")
anova(m0, m1a, m2, test = "F")

cat("\n\n--------------------------------------------\n")
cat("Null model -> Year -> Region + Year \n")
cat("--------------------------------------------\n")
anova(m0, m1b, m2, test = "F")


```

#### Test effect of region and year, 2015-2016 and 2021 only    

```{r}

df_pfos_cod_list <- list(
  df_pfos_cod %>% filter(Year %in% 2015:2016),
  df_pfos_cod %>% filter(Year %in% 2021)
)

names(df_pfos_cod_list) <- c("2015-2016", "2021")

map_int(df_pfos_cod_list, nrow)

for (i in 1:2){
  
  cat("\n==========\n", names(df_pfos_cod_list)[i], "\n==========\n")
  
  m0 <- lm(log_Conc_EQS_ratio ~ 1, data = df_pfos_cod_list[[i]])
  m1a <- lm(log_Conc_EQS_ratio ~ MSFD_regionname, data = df_pfos_cod_list[[i]])
  anova(m0, m1a, test = "F") %>% print()

}


```


#### Region effect, post-hoc test  

```{r}

model <- aov(log_Conc_EQS_ratio ~ MSFD_regionname, data = df_pfos_cod)
multcomp <-  HSD.test(model, "MSFD_regionname", alpha = 0.05)
plot(multcomp)
# par(mfrow=c(1,1), mar=c(3,10,2,3))
# plot(multcomp, horiz=TRUE, las = 1, cex.names = 0.8, xlim = c(-1,0))

```


#### Region effect, post-hoc test, 2015-2016 and 2021 only  

```{r}


for (i in 1:2){
  
  # cat("\n==========\n", names(df_pfos_cod_list)[i], "\n==========\n")
  
  model <- aov(log_Conc_EQS_ratio ~ MSFD_regionname, data = df_pfos_cod_list[[i]])
  multcomp <-  HSD.test(model, "MSFD_regionname", alpha = 0.05)
  plot(multcomp, main = names(df_pfos_cod_list)[i])
  
}

```


### PFOS in blue mussel  

```{r}

species <- "Blue mussel"

gg <- df_pfos %>%
  filter(Common_name == species) %>%
  ggplot(aes(x = MSFD_regionname, y = Conc_EQS_ratio, color = MSFD_regionname)) +
  geom_hline(yintercept = 1) +
  geom_boxplot() +
  scale_color_manual(values = region_colors) +
  scale_y_log10(labels = scales::label_number()) +
  theme_bw() +
  easy_rotate_labels("x", angle = -45) +
  easy_remove_axes(which = "x", what = "title") +
  labs(y = "Ratio concentration/EQS",
       title = paste("PFOS in", tolower(species)))


fn <- glue("Figures/2025/32_PFAS_EQSratio_{tolower(species)}.png")
if (save_plots)
  ggsave(fn, gg, width = 8, height = 5, dpi = 200)

gg

```


### PFOS in birds    

```{r}

speciesgroup <- "Seabird"

# For adding text att bottom of plot
# 
df_speciesnames <- df_pfos %>%
  filter(Species_type == speciesgroup) %>% # count(MSFD_regionname, Common_name) %>%
  group_by(MSFD_regionname) %>%
  # collapse = "\n" gives a line shift 
  summarize(Common_names = paste(unique(Common_name), collapse = "\n")) %>%
  mutate(y = 0.01)

gg <- df_pfos %>%
  filter(Species_type == speciesgroup) %>% # count(MSFD_regionname, Common_name) %>%
  ggplot(aes(x = MSFD_regionname, y = Conc_EQS_ratio)) +
  geom_hline(yintercept = 1) +
  geom_boxplot(aes(color = MSFD_regionname)) +
  geom_text(data = df_speciesnames, aes(y = y, label = Common_names), vjust = 1, size = 3) +
  scale_color_manual(values = region_colors) +
  # Y limit set by trial and error
  scale_y_log10(labels = scales::label_number(), limits = c(0.005, 100)) +
  theme_bw() +
  easy_rotate_labels("x", angle = -45) +
  easy_remove_axes(which = "x", what = "title") +
  labs(y = "Ratio concentration/EQS",
       title = "PFOS in seabirds")

if (save_plots)
  ggsave("Figures/2025/32_PFAS_EQSratio_seabirds.png", gg, width = 8, height = 5, dpi = 200)

gg

```


### PFOS in seabirds and mammals      

```{r}

speciesgroup <- c("Seabird", "Mammal")
# speciesgroup <- c( "Mammal")

# For adding text att bottom of plot
# 

region2_order <- c("Barents Sea, Seabird", "Atlantic Ocean, Seabird", 
                   "Atlantic Ocean, Mammal", "Greater North Sea, Seabird", "Baltic Sea, Seabird")

df_speciesnames <- df_pfos %>%
  filter(Species_type %in% speciesgroup) %>% # count(MSFD_regionname, Common_name) %>%
  mutate(
    MSFD_regionname2 = paste0(MSFD_regionname, ", ", Species_type),
    MSFD_regionname2 = factor(MSFD_regionname2, levels = region2_order)) %>%
  group_by(MSFD_regionname, MSFD_regionname2, Species_type) %>%
  # collapse = "\n" gives a line shift 
  summarize(Common_names = paste(unique(Common_name), collapse = "\n")) %>%
  mutate(y = 0.01)

dput(df_speciesnames$MSFD_regionname2)

gg <- df_pfos %>%
  filter(Species_type %in% speciesgroup) %>% # count(MSFD_regionname, Common_name) %>%
  mutate(
    MSFD_regionname2 = paste0(MSFD_regionname, ", ", Species_type),
    MSFD_regionname2 = factor(MSFD_regionname2, levels = region2_order)) %>%
  ggplot(aes(x = MSFD_regionname2, y = Conc_EQS_ratio)) +
  geom_hline(yintercept = 1) +
  geom_boxplot(aes(color = MSFD_regionname)) +
  geom_text(data = df_speciesnames, aes(y = y, label = Common_names), vjust = 1, size = 3) +
  scale_color_manual(values = region_colors) +
  # Y limit set by trial and error
  scale_y_log10(labels = scales::label_number(), limits = c(0.002, 100)) +
  theme_bw() +
  easy_rotate_labels("x", angle = -45) +
  easy_remove_axes(which = "x", what = "title") +
  labs(y = "Ratio concentration/EQS",
       title = "PFOS in seabirds and mammals")

if (save_plots)
  ggsave("Figures/2025/32_PFAS_EQSratio_seabird_mammals.png", gg, width = 10, height = 5, dpi = 200)

gg

```